---
title: "Week Two: Summarizing & Visualizing Numerical Data"
format:
  html:
    number-sections: true
    number-depth: 2
    section-divs: true
editor: visual
execute: 
  echo: true
  warning: false
  message: false
---

```{r}
#| include: false
#| label: packages

library(tidyverse)
library(gapminder)
library(emo)
library(openintro)
library(unvotes)
library(lubridate)
library(DT)
library(gridExtra)
```

Welcome!

In this week's coursework we are going to continuing exploring data, through data summaries and visualizations, focusing specifically on **numerical** variables. We will be using the **dplyr** package in `R` to wrangle our data and the **ggplot2** package to created data visualizations.

## Learning Outcomes

By the end of this coursework you should be able to:

-   outline the differences between numerical and categorical variables
-   describe what type of summary statistic is appropriate for a given distribution of a numerical variable
-   discuss when it is / is not appropriate to summarize a variable with a mean
-   create visualizations of one and two numerical variables
-   discuss the benefits and shortcomings of different visualizations

# Prepare

## Textbook Reading -- Part 1

::: column-margin
```{r}
#| echo: false
#| out-width: 40%
#| label: picture-of-ims
knitr::include_graphics(here::here("images", 
                                   "ims.jpeg")
                        )
```
:::

[**Required Reading:** Exploring Numerical Data](https://openintro-ims.netlify.app/explore-numerical.html)

### Reading Guide -- Due Tuesday by the start of class

[Download the Word Document](reading-guide/week2-reading2.docx)

::: callout-note
# Submission

Submit your completed reading guide to the Canvas assignment portal!
:::

## Concept Quiz – Due Tuesday by the start of class

::: callout-note
The two concept quizzes from each chapter have been combined into **one** concept quiz on Canvas.
:::

1.  Suppose we have data on the departure delays of flights flying out of New York. What shape would you expect the distribution of departure delays to have?

-   right skew
-   left skew
-   bimodal
-   mutimodal
-   uniform

*Hint:* Think about how you would "typically" expect flight delays to behave.

2.  To better decide what summary statistic we should use to summarize the departure delays it would be best to create a data visualization of the *distribution* of departure delays. What type of visualizations could we make? Select all that apply!

-   boxplot
-   histogram
-   barplot
-   density plot
-   scatterplot

## Textbook Reading -- Part 2

[**Required Reading:** Data Visualization](https://moderndive.com/2-viz.html)

::: column-margin
```{r}
#| echo: false
#| out-width: 60%
#| label: picture-of-modern-dive

knitr::include_graphics(here::here("images", 
                                   "modern_dive.png")
                        )
```
:::

### Reading Guide -- Due Wednesday by the start of class

[Download the Word Document](reading-guide/week2-reading1.docx)

::: callout-note
# Submission

Submit your completed reading guide to the Canvas assignment portal!
:::

## Concept Quiz -- Due Wednesday by the start of class

::: callout-note
The two concept quizzes from each chapter have been combined into **one** concept quiz on Canvas.
:::

1.  What **aesthetics** are being used in the following plot?

*Hint: Think about what goes inside of the `aes()` function and what does not.*

```{r}
#| echo: false
#| message: false
#| warning: false
#| out-width: 70%
#| fig-align: center
#| label: un-votes-facets-colors


un_votes %>%
  filter(country %in% c("United States", "Turkey")) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  inner_join(un_roll_call_issues, by = "rcid") %>%
  mutate(issue = as.character(issue),
         issue = case_when(
           issue == "Nuclear weapons and nuclear material" ~ 
             "Nuclear weapons and \nmaterials",
           issue == "Arms control and disarmament" ~ 
             "Arms control and \ndisarmament", 
           TRUE ~ issue)
         ) %>%
  group_by(country, year = year(date), issue) %>%
  summarize(
    votes = n(),
    percent_yes = mean(vote == "yes")
    ) %>%
  filter(votes > 5) %>%  # only use records where there are more than 5 votes
  mutate(percent_yes = percent_yes * 100) %>% 
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE) +
    facet_wrap(~ issue) +
    labs(
      title = "Percentage of 'Yes' votes in the UN General Assembly",
      y = "",
      x = "Year",
      color = "Country"
    ) + 
  scale_y_continuous(labels = scales::percent_format(scale = 1))

```

-   `x` axis
-   `y` axis
-   `color`
-   `facet`s
-   `point`s
-   `line`s

2.  What **geometric objects** are being used in the displayed visualization?

*Hint: Think about what `geom`s you would use to make this plot!*

-   points
-   lines / smoothers
-   colors
-   facets

3.  What aspects of the distribution of departure delays can you see in the histogram that you could not see in the boxplot?

```{r}
#| echo: false 
#| message: false
#| warning: false
#| layout-nrow: 1

nycflights %>% 
  ggplot(mapping = aes(x = dep_delay)) + 
  geom_boxplot() + 
  scale_x_log10() + 
  labs(x = "Departure Delay (min) -- log transformed")

  
nycflights %>% 
  ggplot(mapping = aes(x = dep_delay)) + 
  geom_histogram(binwidth = 0.5) + 
  scale_x_log10() + 
  labs(x = "Departure Delay (min) -- log transformed")
```

-   shape of distribution
-   median
-   outliers
-   mode

# R Tutorial -- Due Wednesday by the start of class

`r emo::ji("computer")` [**Required Tutorial:** Exploring Numerical Data](https://openintro.shinyapps.io/ims-02-explore-02/)

`r emo::ji("computer")` [**Required Tutorial:** Summarizing with Statistics](https://openintro.shinyapps.io/ims-02-explore-03/)

::: callout-note
# Submission

Submit a screenshot of the completion page for each tutorial to the Canvas assignment portal!
:::
