---
title: "Week Five: Diving Deeper into Regression (Multiple Linear Regression)"
format: 
  html:
    embed-resources: true
    standalone: true
editor: visual
---

```{r}
#| include: false

library(tidyverse)
library(emo)
library(moderndive)
```

Welcome!

In this week's coursework we are going to build off the concepts we learned last week and delve deeper into linear regression. We are going to explore *multiple linear regression*, a statistical model where we have multiple explanatory variables and a single numerical response. We are going to refresh how to visualize these types of models, practice fitting these types of models in `R`, and the learn how to interpret these types of models.

## Learning Outcomes

By the end of this coursework you should be able to:

-   describe to someone what a multiple linear regression is
-   outline how a categorical explanatory variable can be included in a simple linear regression
-   visualize multiple linear regression models with one numerical and one categorical explanatory variable
-   calculate the simple linear regression line for each group of a categorical variable
-   outline how a second numerical explanatory variable can be included in a simple linear regression
-   visualize multiple linear regression models with two numerical explanatory variables
-   interpret the coefficient of each explanatory variable included in the regression model
-   recite different methods that can be used to decide what multiple linear regression model is "best"
-   describe the benefits and costs of using "model selection" for deciding on a multiple linear regression model

# Prepare

## Textbook Reading

::: columns
::: {.column width="40%"}
</br>

```{r}
#| echo: false
knitr::include_graphics(here::here("images", 
                                   "modern_dive.png")
                        )
```
:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
</br> [Chapter 6 (https://moderndive.com/6-multiple-regression.html)](https://moderndive.com/6-multiple-regression.html)
:::
:::

### Reading Guide -- Due Tuesday by noon

[Download the Word Document](reading-guide/week5-reading.docx)

## Concept Quiz -- Due Tuesday by noon

```{r}
#| layout-nrow: 1
#| echo: false
#| message: false
#| warning: false

evals %>% 
  ggplot(mapping = aes(x = age, y = score, color = ethnicity)) +
  geom_jitter(width = 0.2, height = 0.2, alpha = 0.75) + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x = "Age of Professor", 
       y = "Evaluation Score", 
       color = "Ethnicity", 
       title = "Interaction Model") + 
  theme_bw()

evals %>% 
  ggplot(mapping = aes(x = age, y = score, color = ethnicity)) +
  geom_jitter(width = 0.2, height = 0.2, alpha = 0.75) + 
  geom_parallel_slopes(method = "lm", se = FALSE) + 
  labs(x = "Age of Professor", 
       y = "Evaluation Score", 
       color = "Ethnicity", 
       title = "Interaction Model") + 
  theme_bw()
```

**Question 1** -- Based on the visualizations above, I believe the \[interaction / parallel slopes\] model is more appropriate because the slopes between the groups are \[very different / very similar\].

**Question 2** -- What type of model does the following code obtain?

```{r}
minority_lm <- lm(score ~ age * ethnicity, data = evals)
```

-   interaction model
-   parallel slopes model
-   simple linear regression model

</br>

The following is the output from the above regression model (from Question 2):

```{r}
#| echo: false
get_regression_table(minority_lm)
```

</br>

**Question 3** -- The `intercept` line represents the \[evaluation score / mean evaluation score\] for \[male / female / minority / non-minority\] faculty.

</br>

**Question 4** -- The `age` line represents the relationship between age and evaluation scores for \[male / female / minority / non-minority\] faculty.

</br>

**Question 5** -- The `ethnicity:not minority` line represents the \[mean / adjustment to the mean\] evaluation score for \[male / female / minority / non-minority\] faculty.

</br>

**Question 6** -- The `age:ethnicitynot minority` line represents the \[slope / adjustment to the slope\] for the relationship between age and evaluation scores for \[male / female / minority / non-minority\] faculty.

</br>

**Question 7** -- The value of the `age:ethnicitynot minority` line (-0.004) \[does / does not \] match the decision I made in Question 1 as there is \[little difference / substantial difference\] in the slopes between the minority and non-minority faculty.

</br>

**Question 8** -- What type of model does the following code obtain?

```{r}
bty_lm <- lm(score ~ age + bty_avg, data = evals)
```

-   multiple linear regression with two numerical predictors
-   interaction model
-   parallel slopes

</br>

The following is the output from the above regression model (from Question 8):

```{r}
#| echo: false
get_regression_table(bty_lm)
```

</br>

**Question 9** -- The intercept represents the \[\] for professors whose age is \_\_ and who have a average beauty score of \_\_\_.

</br>

**Question 10** -- We interpret the value of -0.003 by age as:

For every \[1 day / 1 year / 1 evaluation\] increase in professor's \[evaluation score / age / average beauty\] we expect the \[course evaluation score / mean course evaluation score\] to \[increase / decrease\] by \_\_, after accounting for \[ethnicity / gender / average beauty scores\].

## R Tutorial -- Due Thursday by noon

1.  [Regression modeling: Parallel Slopes](https://openintro.shinyapps.io/ims-03-model-06/)

2.  [Regression modeling: Evaluating and extending parallel slopes model](https://openintro.shinyapps.io/ims-03-model-07/#section-evaluating-and-extending-parallel-slopes-model)
