---
title: "Week Four: Introduction to Linear Regression (Basic Regression)"
format: html
editor: visual
embed-resources: true
standalone: true
---

```{r}
#| include: false

library(gapminder)
library(emo)
library(openintro)
library(moderndive)
library(skimr)
```

Welcome!

In this week's coursework we are going to start exploring statistical methods---linear regression. We will start with the "basic" case or simple linear regression, with a single quantitative explanatory variable and a quantitative response. We will review the concepts behind linear regression, learn how to visualize regression models, and practice obtaining the estimated regression lines in `R`.

## Learning Outcomes

By the end of this coursework you should be able to:

-   describe the difference between explanatory and predictive modeling
-   produce data visualizations for a simple linear regression
-   describe how `R` decides on the "best" regression line
-   fit a simple linear regression in `R`
-   write out an estimated regression line
-   interpret the slope of an estimated regression line
-   interpret the intercept of an estimated regression line
-   describe what an "offset" is in a simple linear regression
-   interpret the offsets of an estimated regression line
-   calculate a residual for an observation

# Prepare 

## Textbook Reading

::: columns
::: {.column width="40%"}
```{r}
#| echo: false
knitr::include_graphics(here::here("images", 
                                   "modern_dive.png")
                        )
```
:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
</br> [Chapter 5 (https://moderndive.com/5-regression.html)](https://moderndive.com/5-regression.html)
:::
:::

::: callout-note
Don't worry about Section 5.3.3, it goes a bit too in the weeds for what we are interested in.
:::

### Reading Guide

[Download the Word Document](reading-guide/week3-reading.docx)

::: {.callout-note}
# Submission

Submit your completed reading guide to the Canvas assignment portal!
:::

## Concept Quiz -- Due Tuesday by noon

**1. What does the correlation coefficient measure?**

-   strength of a relationship between two variables
-   strength of a linear relationship between two variables
-   strength of a relationship between two numerical variables
-   strength of a linear relationship between two numerical variables

**2. How do you interpret a slope coefficient?**

-   For any increase of $x$, the slope is the expected increase in $y$
-   For any increase of $x$, the slope is the expected increase in the mean of $y$
-   For a 1-unit increase in $x$, the slope is the expected increase in $y$
-   For a 1-unit increase in $x$, the slope is the expected increase in the mean of $y$

**3. How do you interpret an intercept coefficient?**

-   For a 1-unit increase in $x$, the intercept is the expected increase in $y$
-   For a 1-unit increase in $x$, the intercept is the expected increase in the mean of $y$
-   For an $x$ value of 0, the intercept is the expected value of $y$
-   For an $x$ value of 0, the intercept is the expected value of the mean of $y$

**4. Match the *explanatory* and *response* variables to the correct variables in the `lm()` function syntax.**

`lm(variable1 ~ variable2, data = <NAME OF DATASET>)`

**5. What `geom_` adds a linear regression line to a scatterplot?**

-   `geom_line()`
-   `geom_smoother()`
-   `geom_regression()`
-   `geom_smooth()`

## `r emo::ji("computer")` R Tutorial -- Due Thursday by noon

1.  [Regression modeling: Simple linear regression](https://openintro.shinyapps.io/ims-03-model-03/)

2.  [Regression modeling: Interpreting regression models](https://openintro.shinyapps.io/ims-03-model-04/)
