---
title: "The Flaws of Averages"
format: 
  revealjs:
    theme: style.scss
editor: visual
execute: 
  echo: false
---

# Lab 1

## Revisions

::: columns
::: {.column width="47%"}
::: {style="font-size: 1.25em;"}
"Complete" = Satisfactory
:::

-   Your images were included in the document
-   You provided responses to every question
:::

::: {.column width="5%"}
:::

::: {.column width="47%"}
::: {style="font-size: 1.25em;"}
"Incomplete" = Growing
:::

-   Your images [were not]{.underline} included in the document
-   You [did not]{.underline} provide responses to every question
:::
:::

## Key - Code Chunk Options

A code chunk option is declared after a `#|`. Here are some options we may want to use:

::: small
-   `#| label: packages` -- creates a label for the code chunk (describing its contents)
-   `#| echo: false` -- tells Quarto not to output the code in the rendered HTML (only the output)
-   `#| include: false` -- tells Quarto not to include the code or the output in the rendered HTML
:::

## Key - Previewing Your Data

The `glimpse()` function is a great tool to preview the dataset you are working with! It gives you:

-   the dimensions of the data (rows and columns)
-   the names of the columns
-   the data type of each column (e.g., `chr`, `dbl`)
-   a preview of the first 10 rows of each column

## Key - Plotting Your Data

Now that we've practiced making some plots, we know...

::: small
`mapping = aes(y = manufacturer, x = hwy)` declares what variables are plotted on the x- and y-axis.

::: callout-tip
The variable names you put insides `aes()` must be [**identical**]{.underline} to the names of the variables in the dataset!
:::
:::

. . .

</br>

::: small
`labs(x = "Highway Miles Per Gallon", y = "Car Manufacturer")` declares new x- and y-axis labels for the plot.

::: callout-tip
Including nice axis labels (with their units) is a critical part of [**every**]{.underline} visualization we make!
:::
:::

## Completing Revisions

::: {style="font-size: 1.25em;"}
Lab 1 revisions are due by Wednesday, April 17 (at midnight).
:::

::: small
1.  Read comments on Canvas
2.  Go back into your Lab 1 on Posit Cloud and complete your revisions
3.  Render your revised Lab 1
4.  Download your revised HTML
:::

::: small
::: callout-warning
# Reflections

Revisions are required to be accompanied with reflections on what you learned while completing your revisions. These can be written in your Lab 1 Quarto file (next to the problems you revised), in a Word document, or in the comment box on Canvas.
:::
:::

## 

::: {style="font-size: 3em; color: #ed8402;"}
15-minutes
:::

::: {style="font-size: 1.5em;"}
-   Review Lab 1 comments
-   Ask questions
-   Start revisions
:::

##  {background-color="#abdbe3"}

```{r packages}
library(tidyverse)
library(kableExtra)
```

::: {style="font-size: 2em; color: #000000;"}
Suppose...
:::

> ["Overall this instructor was educationally effective."]{style="color: #ed8402;"}

. . .

```{r}
#| echo: false
#| label:  table-of-eval-scores

evals <- tibble( 
  year = c(rep(2021, 2), rep(2022, 5) ), 
  quarter = c(rep("Fall", 2), 
              rep("Winter", 2), 
              rep("Spring", 3) 
              ),
  average = c(4.53, 4.36, 4.18, 4.24, 4.83, 4.41, 4) 
  )

kable(evals) %>% 
  kable_styling(stripe_color = "light blue")
```

## 

<center>

::: {style="font-size: 3.5em; color: #ed8402;"}
How were those averages calculated?
:::

</center>

## 

<center>

::: {style="font-size: 3.5em; color: #ed8402;"}
What do these averages mean?
:::

</center>

## 

::: {style="font-size: 2.5em; color: #000000;"}
The Problem
:::

</br>

It's incredibly rare for scientists, including statisticians, to explicitly think about that conditions underlying their models.

. . .

::: small
> "I've had many conversations in very different contexts with scientists about what the average calculated from the data (or mean in a model) could reasonably represent and whether that was really what the scientist was after." [Dr. Megan Higgs](https://critical-inference.com/more-on-average/)
:::

## 

::: {style="font-size: 2.5em; color: #000000;"}
Why so much resistance?
:::

</br>

. . .

::: {style="font-size: 0.75em; color: #000000;"}
Departments hold specific expectations of statistics courses

</br>

::: fragment
These expectations are conditional on the assumption that means represent the magic quantity of interest.
:::

</br>

::: fragment
I'm then expected to educate you to "play the game" in the scientific culture of averages
:::
:::

## 

::: {style="font-size: 3.5em; color: #000000;"}
Averagarianism
:::

::: {style="font-size: 0.75em; color: #000000;"}
> "The primary research method of averagarianism is [aggregate, then analyze]{style="color: #ed8402;"}: First, [combine many people together and look for patterns in the group]{style="color: #ed8402;"}. Then, use these group patterns (such as averages and other statistics) to analyze and model individuals. The science of the individual instead instructs scientists to analyze, then aggregate: First, look for pattern within each individual. Then, look for ways to combine these individual patterns into collective insight."
>
> The End of Average by Todd Rose
:::

## 

::: {style="font-size: 2.5em; color: #000000;"}
"We've always done it this way"
:::

::: {style="font-size: 0.75em; color: #000000;"}
Methods based on averages are available, easy, convenient, and take little creativity --- and they are expected in our scientific culture.

</br>

::: fragment
Justification for using averages is simply not demanded --- though justification for use of anything but averages is incredibly difficult to sell.
:::
:::

##  {background-color="#f48153"}

::: {style="font-size: 3.5em; color: #000000;"}
Some Rules to Play By
:::

</br>

::: {style="font-size: 0.75em; color: #000000;"}
::: incremental
-   Look at and understand your *raw* data before aggregating
-   Boxplots don't count as visualizing the *raw* data
:::
:::

# Lab 2

<!-- ##  {background-color="#B6CADA"} -->

<!-- ::: {style="font-size: 2em; color: #000000;"} -->

<!-- How Smart are You? -->

<!-- ::: -->

<!-- ::: columns -->

<!-- ::: {.column width="30%"} -->

<!-- (2 minutes) -->

<!-- -   Write two *criteria* would you use to *rank* yourself compared to everyone else in this class -->

<!-- ::: -->

<!-- ::: {.column width="2%"} -->

<!-- ::: -->

<!-- ::: {.column width="30%"} -->

<!-- (3 minutes) -->

<!-- -   Talk with the person on your right about the criteria you proposed -->

<!-- ::: -->

<!-- ::: {.column width="2%"} -->

<!-- ::: -->

<!-- ::: {.column width="30%"} -->

<!-- (5 minutes) -->

<!-- -   Share out -->

<!-- -   Person with most vowels in name should be prepared to share! -->

<!-- ::: -->

<!-- ::: -->

## Departure Delays

-   Inspect the `nycflights` dataset
-   Visualize departure delays
-   Play with histogram `binwidth`
-   `filter()` data to include only certain flights
-   `calculate()` summary statistics
-   Make decisions based on summary statistics
-   Compare summary statistics to a visualization

# Working in Groups

## 

::: {style="font-size: 2em; color: #000000;"}
Cooperative Learning
:::

is a structured form of small-group learning

. . .

</br>

::: {style="font-size: 0.85em;"}
-   Roles relate to **how** the work should be done
    -   Roles **are not** about breaking up the work intellectually
-   Roles allow each person to contribute to the group in significant ways
    -   Each person's participation is necessary to complete the task
:::

##  {background-color="#B6CADA"}

::: {style="font-size: 3em; color: #000000;"}
Group Norms
:::

::: columns
::: {.column width="45%"}
-   Zero tolerance for: racism, sexism, homophobia, transphobia, ageism, ableism

-   Respect one another

-   Intent and impact *both* matter
:::

::: {.column width="5%"}
:::

::: {.column width="45%"}
-   Non-judgmental

-   Take space, make space

-   Embrace discomfort

-   Make decisions by consensus
:::
:::

## 

::: {style="font-size: 2em; color: #000000;"}
How We Learn Together
:::

::: columns
::: {.column width="45%"}
::: {style="font-size: 0.85em;"}
-   No one is done until everyone is done

-   You have the *right* to ask **anyone** in your group for help

-   You have the *duty* to help **anyone** in your group who asks for help
:::
:::

::: {.column width="5%"}
:::

::: {.column width="45%"}
::: {style="font-size: 0.85em;"}
-   Helping someone means explaining your thinking **not** giving answers or doing the work for them

-   Provide a justification when you make a statement

-   Think and work together -- don't divide up the work
:::
:::
:::

## If you chose to participate in the group collaboration research project...

Please fill out the [post-collaboration survey](https://calpolykinesiology.az1.qualtrics.com/jfe/form/SV_4T1tIgzjg6WUlXE)!
