---
title: "The Flaws of Averages"
format: 
  revealjs:
    theme: style.scss
editor: visual
execute: 
  echo: false
---

##  {background-color="#abdbe3"}

```{r packages}
library(tidyverse)
library(kableExtra)
```

::: {style="font-size: 2em; color: #000000;"}
Suppose...
:::

> ["Overall this instructor was educationally effective."]{style="color: #ed8402;"}

. . .

```{r}
#| echo: false
#| label:  table-of-eval-scores

evals <- tibble( 
  year = c(rep(2021, 2), rep(2022, 5) ), 
  quarter = c(rep("Fall", 2), 
              rep("Winter", 2), 
              rep("Spring", 3) 
              ),
  average = c(4.53, 4.36, 4.18, 4.24, 4.83, 4.41, 4) 
  )

kable(evals) %>% 
  kable_styling(stripe_color = "light blue")
```

##  

<center>

::: {style="font-size: 3.5em; color: #ed8402;"}
How were those averages calculated?
:::

</center>

##  

<center>

::: {style="font-size: 3.5em; color: #ed8402;"}
What do these averages mean?
:::

</center>

## 

::: {style="font-size: 2.5em; color: #000000;"}
The Problem
:::

</br> 

It's incredibly rare for scientists, including statisticians, to explicitly think about that conditions underlying their models.

. . .

::: {.small}
> "I've had many conversations in very different contexts with scientists about what the average calculated from the data (or mean in a model) could reasonably represent and whether that was really what the scientist was after." [Dr. Megan Higgs](https://critical-inference.com/more-on-average/)
:::

## 

::: {style="font-size: 2.5em; color: #000000;"}
Why so much resistance?
:::

</br>

. . .

::: {style="font-size: 0.75em; color: #000000;"}
Departments hold specific expectations of statistics courses

</br>

:::{.fragment}
These expectations are conditional on the assumption that means represent the magic quantity of interest.
:::

</br>

:::{.fragment}
I'm then expected to educate you to "play the game" in the scientific culture of averages
:::
:::

## 

::: {style="font-size: 3.5em; color: #000000;"}
Averagarianism
:::

::: {style="font-size: 0.75em; color: #000000;"}
> "The primary research method of averagarianism is [aggregate, then analyze]{style="color: #ed8402;"}: First, [combine many people together and look for patterns in the group]{style="color: #ed8402;"}. Then, use these group patterns (such as averages and other statistics) to analyze and model individuals. The science of the individual instead instructs scientists to analyze, then aggregate: First, look for pattern within each individual. Then, look for ways to combine these individual patterns into collective insight."
>
> The End of Average by Todd Rose
:::

## 

::: {style="font-size: 2.5em; color: #000000;"}
"We've always done it this way"
:::

::: {style="font-size: 0.75em; color: #000000;"}
Methods based on averages are available, easy, convenient, and take little creativity --- and they are expected in our scientific culture.

</br>

:::{.fragment}
Justification for using averages is simply not demanded --- though justification for use of anything but averages is incredibly difficult to sell.
:::
:::

##  {background-color="#f48153"}

::: {style="font-size: 3.5em; color: #000000;"}
Some Rules to Play By
:::

</br>

::: {style="font-size: 0.75em; color: #000000;"}
::: incremental
-   Look at and understand your *raw* data before aggregating
-   Boxplots (and other plots of aggregated data) don't count as visualizing the *raw* data
:::
:::

# Lab 2

## Departure Delays

-   Inspect the `nycflights` dataset
-   Visualize departure delays
-   Play with `binwidth`
-   Filter data to include only certain flights
-   Calculate summary statistics
-   Make decisions based on summary statistics 
-   Compare summary statistics with a visualization

## If you chose to participate in the group collaboration research project...

Please fill out the [post-collaboration survey](https://calpolykinesiology.az1.qualtrics.com/jfe/form/SV_4T1tIgzjg6WUlXE)! 
