---
title: "Introduction to Multiple Linear Regression"
format: 
  revealjs:
    theme: style.scss
editor: visual
---

```{r set-up}
#| include: false

library(tidyverse)
library(moderndive)
library(openintro)

my_theme <- theme_bw() + 
  theme(axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20), 
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 20),
        panel.border = element_blank(), 
        panel.grid.minor = element_blank()
        )
```

# Midterm Project Week

## Plan for Today

1.  Review different types of multiple linear regression models

2.  Complete an activity on sample selection

3.  Start Midterm Project write-up

## Plan for Wednesday

**No lab** -- focus on getting **all** the coding accomplished for the Midterm Project

## Draft Due by Sunday

To get everyone feedback on their drafts in a timely manner, the first drafts are due by Sunday.

::: callout-caution
# Deadline Extension

A deadline extension [**is**]{.underline} permitted for the first draft. Deadline extensions [**are not**]{.underline} permitted for the final version (due next week).
:::

## Reminders About Deadlines

-   Midterm Project Proposal is / was due [today]{.underline} by 5pm
-   Lab 3 Revisions are due Wednesday (May 1)
-   Second round of Lab 2 revisions are due by Wednesday (May 1)

# Multiple Linear Regression

##  {background-color="#D4D4D4"}

::: {style="font-size: 3em; color: #000000;"}
Before...
:::

```{r old-slr}
#| echo: false
ncbirths %>% 
  ggplot(mapping = aes(x = weeks, y = weight)) + 
  geom_jitter() + 
  geom_smooth(method = "lm") +
  labs(x = "Weeks of Gestation", 
       y = "Weight of Baby (lbs)") +
  my_theme

```

##  {background-color="#D4D4D4"}

::: {style="font-size: 3em; color: #000000;"}
Now...
:::

```{r two-mlr}
#| echo: false

ncbirths %>% 
  ggplot(mapping = aes(x = weeks, y = weight, color = habit)) + 
  geom_jitter(alpha = 0.8) + 
  geom_smooth(method = "lm") +
  labs(x = "Weeks of Gestation", 
       y = "Weight of Baby (lbs)", 
       color = "Parent Smoking Status") + 
  my_theme

```

##  {background-color="#B6CADA"}

</br> </br>

<center>

::: {style="font-size: 6em; color: #000000;"}
How?
:::

</center>

##  {background-color="#D4D4D4"}

::: {style="font-size: 2em; color: #000000;"}
Offsets!
:::

```{r smoke-mlr}
#| eval: false
#| echo: true
#| code-line-numbers: false

smoke_lm <- lm(weight ~ weeks * habit, data = ncbirths)

get_regression_table(smoke_lm)

```

```{r smoke-mlr-output}
#| echo: false
smoke_lm <- lm(weight ~ weeks * habit, data = ncbirths)

get_regression_table(smoke_lm) %>% 
  select(term:std_error)
```

::: callout-note
# Interaction Model

The \* means the variables are interacting!
:::

## Estimated Regression Equations

```{r}
#| echo: false

get_regression_table(smoke_lm) %>% 
  select(term:std_error)
```

::: {style="font-size: 1.1em; color: #0F4C81;"}
What is the regression equation for non-smoker mothers?
:::

. . .

::: {style="font-size: 1.1em; color: #0F4C81;"}
What is the regression equation for smoker mothers?
:::

##  {background-color="#B6CADA"}

<center>

::: {style="font-size: 3em; color: #000000;"}
What if we have a second numerical explanatory variable?
:::

</center>

##  {background-color="#D4D4D4"}

```{r mlr-two-num}
ncbirths %>% 
  ggplot(mapping = aes(x = weeks, y = weight, color = mage)) + 
  geom_jitter() + 
  geom_smooth(method = "lm") +
  labs(x = "Weeks of Gestation", 
       y = "Weight of Baby (lbs)", 
       color = "Mother's Age") +
  my_theme

```

##  {background-color="#D4D4D4"}

::: {style="font-size: 2em; color: #000000;"}
Multiple slopes
:::

```{r age-mlr}
#| eval: false
#| echo: true
#| code-line-numbers: false

age_lm <- lm(weight ~ weeks + mage, data = ncbirths)

get_regression_table(age_lm)

```

```{r age-mlr-output}
#| echo: false
age_lm <- lm(weight ~ weeks + mage, data = ncbirths)

get_regression_table(age_lm) %>% 
  select(term:std_error)
```

. . .

::: {style="font-size: 1.25em; color: #0F4C81;"}
How do you interpret the value of 0.346?
:::

. . .

::: {style="font-size: 1.25em; color: #0F4C81;"}
How do you interpret the value of 0.02?
:::

##  {background-color="#B6CADA"}

<center>

::: {style="font-size: 2.5em; color: #000000;"}
But how do we decide if the interaction model is "best" without a p-value??????
:::

</center>

##  {background-color="#D4D4D4"}

::: {style="font-size: 2em; color: #000000;"}
When investigating if a relationship differs...
:::

. . .

::: {style="font-size: 1.25em; color: #0F4C81;"}
Always start with the "interaction" / different slopes model.
:::

. . .

::: {style="font-size: 1.25em; color: #b76352;"}
If the slopes look different, you're done!
:::

. . .

::: {style="font-size: 1.25em; color: #34605f;"}
If the slopes look similar, then fit the "additive" / parallel slopes model.
:::

## 

::: {style="font-size: 2em; color: #000000;"}
Different Enough?
:::

```{r}
#| echo: false

ggplot(data = MA_schools, 
       mapping = aes(y = average_sat_math, 
                       x = perc_disadvan, 
                       color = size)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  labs(x = "Percent Economically Disadvantaged", 
       y = "Average SAT Math", 
       color = "Size of School") +
  my_theme
```

##  {background-color="#B6CADA"}

::: {style="font-size: 1.5em; color: #000000;"}
What if they're not very different?
:::

```{r}
#| fig-align: center

MA_schools %>% 
  ggplot(mapping = aes(y = average_sat_math, 
                       x = perc_disadvan, 
                       color = size)) + 
  geom_point() +
  geom_parallel_slopes() + 
  labs(x = "Percent Economically Disadvantaged", 
       y = "Average SAT Math", 
       color = "Size of School") +
  my_theme
```

## 

::: {style="font-size: 2em; color: #000000;"}
Parallel Slopes
:::

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false

lm(average_sat_math ~ perc_disadvan + size, 
   data = MA_schools)
```

</br>

```{r}
lm(average_sat_math ~ perc_disadvan + size, 
   data = MA_schools) %>% 
  get_regression_table() %>% 
  select(term:std_error)
```

##  {background-color="#B6CADA"}

::: {style="font-size: 2em; color: #000000;"}
Group equations -- Baseline
:::

```{r}
lm(average_sat_math ~ perc_disadvan + size, data = MA_schools) %>% 
  get_regression_table() %>% 
  select(term:std_error)
```

. . .

$$\widehat{SAT}_{small} = 588 - 2.78 \times \text{percent disadvantaged}$$

##  {background-color="#B6CADA"}

::: {style="font-size: 2em; color: #000000;"}
Group equations -- Offsets
:::

```{r}
lm(average_sat_math ~ perc_disadvan + size, data = MA_schools) %>% 
  get_regression_table() %>% 
  select(term:std_error)
```

. . .

::: small
$$\widehat{SAT}_{medium} = (588 - 11.9) - 2.78 \times \text{percent disadvan}$$
:::

::: small
$$\widehat{SAT}_{medium} = 576.1 - 2.78 \times \text{percent disadvan}$$
:::

. . .

::: small
$$\widehat{SAT}_{large} = (588 - 6.36) - 2.78 \times \text{percent disadvan}$$
:::

::: small
$$\widehat{SAT}_{large} = 581.64 - 2.78 \times \text{percent disadvan}$$
:::

# Sample Selection Activity

## Find Your Data Group!

Once you have found other students working on the same dataset, complete [the sample selection activity](https://atheobold.github.io/stat-313-website/activity/week-5-sample-selection.html).

# Midterm Project Work Time

## Steps Before Wednesday

1.  Insert the description of your dataset and variables (from the Midterm Proposal) into the "Introduction" of your project

2.  Pose a research question about your selected variables, **which can be addressed with multiple linear regression**

3.  Insert the code to create the required two (or three) visualizations

4.  Write a description of what you see in the visualizations

5.  Make a decision which model you believe is "best"

<!-- ## Steps Before Wednesday (STAT 513) -->

<!-- ::: small -->
<!-- 1.  Create a [nicely formatted version of your dataset](../project/data-formatting-513.qmd) -->

<!-- 2.  Insert the description of your dataset and variables (from the Midterm Proposal) into the "Introduction" of your project -->

<!-- 3.  Pose a research question about your selected variables, **which can be addressed with multiple linear regression** -->

<!-- 4.  Work with Dr. Theobold to write code to read in your data! -->

<!-- 5.  Insert the code to create the required two (or three) visualizations -->

<!-- 6.  Write a description of what you see in the visualizations -->

<!-- 7.  Make a decision which model you believe is "best" -->
<!-- ::: -->
