---
title: "Incorporating Categorical Variables"
format: 
  revealjs:
    logo: images/dplyr.png
    theme: night
    embed-resources: true
    standalone: true
editor: visual
execute: 
  echo: false
---

```{r set-up}
#| include: false

library(lterdatasampler)
library(tidyverse)

vertebrates_full_code <- and_vertebrates %>% 
  mutate(section = if_else(section == "CC", 
                           "clear cut forest",
                           "upstream old growth coniferous forest"
                           )
         ) 

trout <- vertebrates_full_code %>% 
  filter(species == "Cutthroat trout")

salamander <- vertebrates_full_code %>% 
  filter(species != "Cutthroat trout")
```


## In `R`...

categorical variables can have either `character` *or* `factor` data types

</br>

. . .

`factor` -- structured & fixed number of levels / options

-   can be ordered or unordered

</br>

. . .

`character` -- unstructured & variable number of levels

-   is inherently unordered

## Incorporating Categorical Variables into Data Visualizations

::: incremental
-   As a variable on the `x`- or `y`-axis

-   As a `color` / `fill`

-   As a facet
:::

## Salamander Size

::: columns
::: {.column width="55%"}
```{r}
#| echo: true
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| code-line-numbers: false

ggplot(data = salamander, 
       mapping = aes(x = length_2_mm)) + 
  geom_histogram(binwidth = 14) + 
  labs(x = "Snout to Tail Length (mm)")
```
:::

::: {.column width="5%"}
:::

::: {.column width="40%"}
::: {style="font-size: 0.9em;"}
</br>

How would this histogram look if there was no variation in salamander length?

</br>

What are possible causes for the variation in salamander length?
:::
:::
:::

## Faceted Histograms

```{r facet-hist}
#| echo: true
#| code-line-numbers: false
#| out-width: 95%

ggplot(data = salamander, 
       mapping = aes(x = length_1_mm)) + 
  geom_histogram(binwidth = 14) + 
  facet_wrap(~ section, scales = "free") +
  labs(x = "Snout to Tail Length (mm)")
```

What do you think `scales = "free"` does?

## Side-by-Side Boxplots

::: columns
::: {.column width="50"}
```{r horizontal-boxplot}
#| fig-width: 6
#| fig-height: 3.5
#| fig-align: center
#| echo: true
#| code-line-numbers: false

ggplot(data = salamander, 
       mapping = aes(x = length_1_mm, 
                     y = species)
         ) + 
  geom_boxplot() + 
  labs(x = "Snout to Tail Length (mm)", 
       y = "Salamander Species") 
```
:::

::: {.column width="50"}
```{r vertical-boxplot}
#| fig-width: 6
#| fig-height: 4
#| fig-align: center
#| echo: true
#| code-line-numbers: false

ggplot(data = salamander, 
       mapping = aes(y = length_1_mm, 
                     x = species)
         ) + 
  geom_boxplot() + 
  labs(y = "Snout to Tail Length (mm)", 
       x = "Salamander Species")

```
:::
:::

Which orientation do you prefer? Vertical or horizontal?

## Colors in Boxplots

```{r boxplot-color}
#| fig-align: center
#| echo: true
#| code-line-numbers: false
#| out-width: 80%

ggplot(data = salamander, 
       mapping = aes(x = length_1_mm, 
                       y = species, 
                       color = unittype)
         ) + 
  geom_boxplot() + 
  labs(x = "Snout to Tail Length (mm)", 
       y = "Salamander Species", 
       color = "Channel Type")
```

::: {style="font-size: 0.5em;"}
Why are there only two boxplots for the Olympic torrent salamander?
:::

## Facets & Colors in Boxplots

```{r boxplot-facets}
#| fig-align: center
#| echo: true
#| code-line-numbers: false
#| output-location: slide

ggplot(data = salamander, 
       mapping = aes(x = length_1_mm, 
                       y = species, 
                       color = section)
         ) + 
  geom_boxplot() + 
  facet_wrap(~ unittype) + 
  labs(x = "Snout to Tail Length (mm)", 
       y = "Salamander Species", 
       color = "Section in Mack Creek")
```

## Facets & Color in Scatterplots

```{r scatterplot-facets}
#| fig-align: center
#| echo: true
#| code-line-numbers: false
#| output-location: slide

ggplot(data = salamander, 
       mapping = aes(x = length_1_mm, 
                       y = weight_g, 
                       color = section)
         ) + 
  geom_point() + 
  facet_wrap(~species, scales = "free") +
  labs(y = "Snout to Tail Length (mm)", 
       x = "Year", 
       color = "Salamander Species")

```

## Your Turn

```{r}
#| echo: false
#| fig-align: center

trout %>%
  filter(unittype %in% c("R", "I", "S")) %>% 
  ggplot(aes(x = length_1_mm, y = weight_g, color = unittype)) + 
  geom_jitter(alpha = 0.5) + 
  labs(x = "Length (mm)", 
       y = "Weight (g)", 
       color = "Channel \nClassification") + 
  theme_light()
```

What are the aesthetics included in this plot?


