---
title: "Midterm Project Work Day"
format: 
  revealjs:
    theme: style.scss
editor: visual
---

# Reminders About Deadlines

##  {background-color="#B6CADA"}

- Lab 3 Revisions are due today (Thursday, February 8)

- Lab 4 Revisions are due next Thursday (February 15)

:::{.callout-caution}
# Making a copy of your group's Lab 4
If you were the recorder (typer) for your group, you need to [make your project public](https://atheobold.github.io/stat-313-website/slides/week4-day2.html#copying-the-lab-last-weeks-recorder). If you were not the recorder, you need to [make a copy of your group's project](https://atheobold.github.io/stat-313-website/slides/week4-day2.html#copying-the-lab-everyone-else). 
:::

- The first draft of your Midterm Project is due on [Sunday at midnight]{.underline}. 

# Comments from Project Proposals

##  For Every Dataset (in STAT 313)

- [Every]{.underine} dataset is included in an R package! You need to find which package that is and load it in!

- If your dataset is associated with a publication, that publication should be referenced in your Introduction!

## Introduction Versus Methods

- The description of your data goes and your research question goes in your Introduction. 

- The description of your variables goes at the beginning of your Methods! 

## Specific Dataset Advice

- For the `and_vertebrates` data, you should include `species` as an explanatory variable. If you don't you are assuming the same relationship applies to trout AND salamanders. 

- For the `hbr_maples` data, you cannot use `year` as a numerical variable. There are only two years of data!

- For the `pie_crabs` data:
  * `size` should be the response variable
  * `site` and `latitude` measure the [same]{.underline} thing

# Coding a Multiple Linear Regression

##  {background-color="#B6CADA"}

::: {style="font-size: 2em; color: #000000;"}
Step 0 -- Read in Your Data
:::

::: columns
::: {.column width="47%"}
STAT 313

:::{.small}
- Locate what package your data live in (found in the directions for the midterm project proposal)

- Load in the package you need! 

- Get started!
:::
:::

::: {.column width="3%"}
:::

::: {.column width="50%"}
STAT 513

:::{.small}
- Upload your formatted dataset to your Midterm Project on Posit Cloud
- Write the following code to load in your dataset:
:::
:::
:::

</br>

```{r}
#| eval: false
#| echo: true
name_you_want <- read_xlsx("name of your dataset.xlsx")
```

**OR**

```{r}
#| eval: false
#| echo: true
name_you_want <- read_csv("name of your dataset.csv")
```

##  {background-color="#B6CADA"}

::: {style="font-size: 2em; color: #000000;"}
Step 1 -- Visualizations
:::

::: columns
::: {.column width="47%"}
[Two Numerical Variables]{.underline}

:::{.callout-tip}
# Three total visualizations
:::

:::{.small}
-   Visualize the model with **both** variables -- `color` gradient
-   Visualize **two** simple linear regression models -- one for each variable
:::
:::

::: {.column width="3%"}
:::

::: {.column width="50%"}
[One Categorical & One Numerical Variable]{.underline}

:::{.callout-tip}
# Two total visualizations
:::

:::{.small}
-   Visualize the model using `geom_smooth(method = "lm")`
-   Visualize the model using `geom_parallel_slopes()`
:::
:::
:::

:::{.callout-warning}
# No `theme_classic()`
I like grid lines! Please use `theme_bw()` instead!
:::

##  {background-color="#D4D4D4"}

::: {style="font-size: 2em; color: #000000;"}
Step 2 -- Decide the "Best" Model
:::

::: columns
::: {.column width="45%"}
[Two Numerical Variables]{.underline}

::: {style="color: #0F4C81; font-size: 0.75em"}
-   If there appears to be a relationship with the colors -- include **both** variables!

-   If the colors are equally dispersed throughout the plot -- choose the [**one**]{.underline} variable that has the [**stronger**]{.underline} relationship (larger slope)!
:::
:::

::: {.column width="5%"}
:::

::: {.column width="45%"}
[One Categorical & One Numerical Variable]{.underline}

::: {style="color: #0F4C81; font-size: 0.75em"}
- Look at the plot where the lines are allowed to be different! Does it look like they are?
:::
:::
:::

##  {background-color="#D4D4D4"}

::: {style="font-size: 2em; color: #000000;"}
Step 3 -- Fit the regression model with `lm()`
:::

::: columns
::: {.column width="45%"}
[Two Numerical Variables]{.underline}

::: {style="color: #0F4C81; font-size: 0.75em"}
-   Are both variables included? Use a `+` to separate them!
:::
:::

::: {.column width="5%"}
:::

::: {.column width="45%"}
[One Categorical & One Numerical Variable]{.underline}

::: {style="color: #0F4C81; font-size: 0.75em"}
-   Are the slopes different? You need to fit a different slopes model! Use a `*` to separate the variables!

-   Are the slopes similar? You need to fit a parallel slopes model! Use a `+` to separate the variables!
:::
:::
:::

##  {background-color="#D4D4D4"}

::: {style="font-size: 2em; color: #000000;"}
Step 4: Get the coefficients with `get_regression_table()`
:::


##  {background-color="#D4D4D4"}

::: {style="font-size: 5em; color: #000000;"}
Now interpret! 
:::


<!-- ##  {background-color="#B6CADA"} -->

<!-- ::: {style="font-size: 2em; color: #000000;"} -->

<!-- **Hermeneutics** hərmə'n(y)o͞odiks -->

<!-- ::: -->

<!-- > *the study of interpretation* -->

<!-- </br> -->

<!-- . . . -->

<!-- > "Philosophically, hermeneutics concerns the meaning of interpretation--its basic nature, scope and validity, and its place within and implications for human existance." -->

<!-- ##  {background-color="#D4D4D4"} -->

<!-- ::: {style="font-size: 2em; color: #0F4C81;"} -->

<!-- **Statistical Hermeneutics** -->

<!-- ::: -->

<!-- > "The study of the interpretation of results and conclusions based on statistical theory, methods, and reasoning." -->

<!-- . . . -->

<!-- -   The complexities in interpretations of statistical models, results, etc. are often downplayed and taken for granted. -->

<!-- -   It is not often that we are hit with an *explicit* call to discuss our *implicit* interpretations of the statements that arise from our use of statistical methods and results. -->

<!-- ##  {background-color="#D4D4D4"} -->

<!-- ::: {style="font-size: 2em; color: #0F4C81;"} -->

<!-- Better for Science -->

<!-- ::: -->

<!-- > "Modern science despite all of the methodological and technological sophistication, has failed to account for the basic epistemic foundation on which it lies." -->

<!-- . . . -->

<!-- -   Hermeneutics possesses a positive attitude toward the finitude of human understanding. -->

<!--     -   The concern is not to eradicate our finitude, but to recognize the consequences of our limits. -->
