---
title: "Connecting ANOVA with Multiple Linear Regression"
format: 
  revealjs:
    theme: style.scss
editor: source
---

```{r}
#| include: false
#| label: set-up

library(tidyverse)
library(palmerpenguins)
library(broom)
library(ggridges)
library(infer)

evals_small <- moderndive::evals %>% 
  group_by(prof_ID) %>% 
  mutate(min_eval = min(score), 
         age_cat = case_when(age < 37 ~ "young faculty",
                             age >= 37 & age < 45 ~ "middle age", 
                             age >= 45 & age < 60 ~ "older faculty", 
                             age >= 60 ~ "nearly retired"),
         age_cat = factor(age_cat, 
                          levels = c("young faculty", "middle age",
                                     "older faculty", "nearly retired")
                          )
         ) %>% 
  distinct(prof_ID, .keep_all = TRUE) 
```

# Lab 8 Recap

## Common Mistakes

# Statistical Critique Recap

## Common Mistakes

## Upcoming Deadlines

-   Lab 7 revisions are due today (by midnight)
-   Lab 8 revisions are due next Thursday
-   Statistical Critique 2 revisions are due next Thursday
-   Final revisions on **all** assignments will be accepted until Sunday, March 17

::: callout-caution
# Don't forget reflections!
:::

# How does this connect with the ANOVA we did in the Statistical Critique?

# 

# Where are we headed next week? 

## 

```{r age-two-cat}
#| echo: false
#| out-width: 70%

ggplot(data = evals_small, 
       mapping = aes(y = age_cat, x = min_eval, fill = gender)) + 
  geom_density_ridges(alpha = 0.5) + 
  labs(y = "Professor's Age", 
       x = "Minimum Evaluation Score", 
       fill = "Professor's Sex") +
  theme(axis.title = element_text(size = 18), 
        axis.text.y = element_text(size = 16), 
        axis.text.x = element_text(size = 14))
```