---
title: "One-Way ANOVA Review"
format: revealjs
editor: visual
---

```{r set-up}
#| include: false
library(tidyverse)
library(lterdatasampler)
library(openintro)
library(moderndive)
library(infer)
library(ggridges)
library(broom)
library(emo)
library(gt)

movies <- movies_sample %>% 
  mutate(era = case_when(year < 1920 ~ "1910s", 
                         year >= 1920 & year < 1930 ~ "1920s", 
                         year >= 1930 & year < 1940 ~ "1930s", 
                         year >= 1940 & year < 1950 ~ "1940s", 
                         year >= 1950 & year < 1960 ~ "1950s", 
                         year >= 1960 & year < 1970 ~ "1960s", 
                         year >= 1970 & year < 1980 ~ "1970s", 
                         year >= 1980 & year < 1990 ~ "1980s", 
                         year >= 1990 & year < 2000 ~ "1990s", 
                         year >= 2000 ~ "2000s")) %>% 
  filter(era %in% c("1970s", "1980s", "1990s", "2000s"))

trout <- filter(and_vertebrates, 
                species == "Cutthroat trout")
```

::: {style="font-size: 2em; color: #000000;"}
Plan for Revisions
:::

::: incremental
- Statistical Critique 2 has been graded
- Lab 8 will be graded by Friday
- You are permitted to submit **one** "free" revision this week
- Additional revisions can be purchased with any tokens you have remaining
- Keep your revisions in mind, if you wish to purchase a 2-day extension on the final project
- Revisions will be accepted until Thursday, June 15
:::

## You've seen a one-way ANOVA model before! 

```{r trout-chanel}
#| echo: true
#| code-line-numbers: false

trout_lm <- lm(weight_g ~ unittype, data = trout)

get_regression_table(trout_lm)
```

</br>

::: fragment
<center>

What do the `estimate`s from each of these lines (e.g., `unittype:IP`) represent?

</center>

:::

## One-Way ANOVA -- Testing if at least one chanel type has a different mean weight

```{r chanel-anova}
#| echo: true
#| code-line-numbers: false

aov(weight_g ~ unittype, data = trout) %>% 
  tidy()
```

. . .

</br>

:::{style="text-align: center;"}

What would you decide?

::: fragment
What conditions need to be checked to know if this p-value is reliable?
:::
:::

## Checking Conditions

```{r channel-}
#| echo: false
#| fig-align: center

trout %>% 
  ggplot(aes(y = unittype, x = weight_g)) + 
  geom_density_ridges(alpha = 0.5, scale = 1) +
  labs(y = "Channel Type", 
       x = "Weight (g)") + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 16), 
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18))
```

How do you feel about the normality condition?

. . .

How do you feel about the equal variance condition?

## Comparing Variances

::: {.columns}
::: {.column width="50%"}
**Un-transformed Variances**

```{r variance-table}
#| echo: false

trout %>% 
  drop_na(weight_g) %>% 
  group_by(unittype) %>% 
  summarize(var = var(weight_g)) %>% 
  gt()
```
:::

::: {.column width="50%"}

::: fragment
**Log Transformed Variances**
```{r variance-table-log}
#| echo: false
trout %>% 
  drop_na(weight_g) %>% 
  group_by(unittype) %>% 
  summarize(var = var(log(weight_g))) %>% 
  gt()
```
:::

:::
:::

. . .

</br>
<center>

What do you think?

</center>

# Two-Way ANOVA Models


##

::: {style="font-size: 3em; color: #000000;"}
Two-way ANOVA
:::

</br>

::: {style="font-size: 2em; color: #0F4C81;"}
Goal:
:::

Assess if [multiple]{style="color: #0F4C81"} categorical variables have a relationship with the response. 


##

::: {style="font-size: 3em; color: #000000;"}
Modeling Options
:::

::: columns
::: {.column width="50%"}
::: {style="font-size: 2em; color: #ed8402;"}
Additive Model
:::

:::{.fragment}
Assess if each explanatory variable has a meaningful relationship with the response, conditional on the variable(s) included in the model. 
:::

:::

::: {.column width="50%"}
::: {style="font-size: 2em; color: #0F4C81;"}
Interaction Model
:::

:::{.fragment}
Assess if the relationship between one categorical explanatory variable and the response **differs** based on the values of another categorical variable. 
:::

:::
:::

##

::: {style="font-size: 2em; color: #000000;"}
**What are we looking for?**
:::

```{r twa-plot}
#| echo: false
#| fig-align: center

movies %>% 
  ggplot(aes(y = era, x = rating, fill = genre)) + 
  geom_density_ridges(alpha = 0.5, scale = 1) +
  labs(y = "Era", 
       x = "IMDB Rating", 
       fill = "Genre") + 
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 16), 
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18))
```

## Another way to think about it...

```{r twa-plot-2}
#| echo: false
#| fig-align: center

movies %>% 
  ggplot(aes(y = era, x = rating)) + 
  geom_density_ridges(alpha = 0.5, scale = 1) +
  labs(y = "Era", 
       x = "IMDB Rating") + 
  facet_wrap(~genre) +
  theme(axis.title = element_text(size = 18), 
        axis.text = element_text(size = 16), 
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18))
```


## Hypothesis Test Steps

::: columns
::: {.column width="35%"}
::: {style="font-size: 1.5em; color: #0F4C81;"}
Step 1: Fit a one-way ANOVA model for each categorical variable
:::
:::

::: {.column width="5%"}
:::

::: {.column width="60%"}
::: fragment
::: {style="font-size: 1.5em; color: #687244;"}
Step 2: Decide if each explanatory variable has a meaningful relationship with the response variable
:::
:::

::: fragment
-   If yes, then go to step 3!
-   If no, then report which variable (if any) has the strongest relationship with the response. 
:::
:::
:::

## Additive Two-Way ANOVA

</br>

```{r}
#| echo: true
#| eval: false
aov(rating ~ era + genre, data = movies) %>% 
  tidy()
```


Why did I use a `+` when fitting the model?

. . .

</br>

```{r}
#| echo: false
aov(rating ~ era + genre, data = movies) %>% 
  tidy() %>% 
  gt() 
```

:::{style="text-align: center;"}

**What is the `era` line testing?**

::: fragment
**What would you decide?**
:::

:::

# Connecting to research questions

## One-way ANOVA Research Question

### 90-seconds

</br>
</br>

How would you phrase a research question for the one-way ANOVA model we've been investigating? 

**Explanatory Variable:** `genre`

**Response Variable:** `rating`

## Two-way ANOVA Research Question

### 90-seconds

</br>
</br>

How would you phrase a research question for the two-way ANOVA model we've been investigating? 

**Explanatory Variables:** `genre` and `era`

**Response Variable:** `rating`

## What about two-way ANOVA interaction models?

::: incremental
- An interaction model is similar to a different slopes multiple linear regression

- It assesses if the relationship between two variables differs when a third variable is considered

- Due to time, we we **are not** fitting interaction models for the Final Project 
:::


