---
title: "🔬 Simulation-Based Methods versus Theory-Based Methods"
format: 
  revealjs:
    theme: dark
editor: visual
---

```{r set-up}
library(tidyverse)
library(infer)
library(moderndive)
library(lterdatasampler)

my_theme <- theme(axis.title.x = element_text(size = 18), 
                  axis.title.y = element_text(size = 18), 
                  axis.text.x = element_text(size = 12), 
                  axis.text.y = element_text(size = 12))

obs_slope <- evals %>% 
  specify(response = score, 
          explanatory = bty_avg) %>% 
  calculate(stat = "slope")
```

# Lab 7 Recap

## Common Mistakes

2) Question 8 (Interpret Confidence Interval): A lot of groups either did not include the population or they stated the population was the 13 marshes and not all marshes on the East coast. For this question, only one or two groups said the population was about crabs!

3) Question 11 (Bootstrap Assumptions): A lot of groups did not talk about marshes and instead mentioned resampling. Some also mentioned that the population was crabs :(


# Thinking about the Final Project...


## 

::: {style="font-size: 4em; color: #FFFFFF;"}
What did we do on Tuesday?
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
We carried out a hypothesis test!
:::

::: columns
::: {.column width="35%"}
$$H_0: \beta_1 = 0$$

$$H_A: \beta_1 \neq 0$$
:::

::: {.column width="5%"}
:::

::: {.column width="60%"}
```{r evals-slr}
#| message: false
#| fig-align: center

ggplot(data = evals, 
       mapping = aes(x = bty_avg, y = score)) +
  geom_jitter() + 
  geom_smooth(method = "lm") +
  labs(x = "Average Beauty Score", 
       y = "Course Evaluation Score") +
  my_theme
```
:::
:::

::: {style="font-size: 1em; color: #ed8402;"}
What do these hypotheses mean *in words*?
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
By creating a permutation distribution!
:::

```{r evals-permute}
#| echo: true
null_dist <- evals %>% 
  specify(response = score, 
          explanatory = bty_avg) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "slope")
```

. . .

</br>

::: {style="font-size: 1.5em; color: #B6CADA;"}
What is happening in the `generate()` step?
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
And visualizing where our observed statistic fell on the distribution
:::

::: columns
::: {.column width="60%"}
```{r}
visualise(null_dist) +
  shade_p_value(obs_stat = obs_slope, direction = "two-sided") +
  labs(x = "Permuted Slope Statistic")
```
:::

::: {.column width="5%"}
:::

::: {.column width="35%"}
::: {style="font-size: 1.5em; color: #B6CADA;"}
What would you estimate the p-value to be?
:::
:::
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
And calculated the p-value
:::

::: columns
::: {.column width="60%"}
```{r}
#| echo: true

get_p_value(null_dist, 
            obs_stat = obs_slope, 
            direction = "two-sided")
```
:::

::: {.column width="5%"}
:::

::: {.column width="35%"}
::: {style="font-size: 1.5em; color: #B6CADA;"}
</br> </br> What would you decide for your hypothesis test?
:::
:::
:::

## 

::: {style="font-size: 2.5em; color: #FFFFFF;"}
How would this process have changed if we used theory-based methods instead?
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
Approximating the permutation distribution
:::

::: columns
::: {.column width="40%"}
A $t$-distribution can be a reasonable approximation for the permutation distribution if certain conditions are not violated.
:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
```{r t-distribution}
plot(c(-5, 5),
     c(0, dnorm(0)),
      type = "n", 
     ylab = "", 
     xlab = "",
     axes = FALSE
     )

at <- seq(-10, 10, 2)
axis(1, at)
axis(1, at - 1, rep("", length(at)), tcl = -0.1)
abline(h = 0)
COL. <- openintro::fadeColor(openintro::IMSCOL["blue", "full"], c("FF", "89", "68", "4C", "33"))
COLt <- openintro::fadeColor(openintro::IMSCOL["blue", "full"], c("FF", "AA", "85", "60", "45"))
DF <- c("normal", 8, 4, 2, 1)
X <- seq(-10, 10, 0.02)
Y <- dnorm(X)
lines(X, Y, col = COL.[1])
for (i in 2:5) {
  Y <- dt(X, as.numeric(DF[i]))
  lines(X, Y, col = COL.[i], lwd = 1.5)
}
legend(2.5, 0.4,
  legend = c(
    DF[1],
    paste("t, df = ", DF[2:5], sep = "")
  ),
  col = COL.,
  text.col = COLt,
  lty = rep(1, 5),
  lwd = 1.5
)

```
:::
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
What about the observed statistic?
:::

::: panel-tabset
## Before

```{r obs-slop}
#| echo: true
obs_slope <- evals %>% 
  specify(response = score, 
          explanatory = bty_avg) %>% 
  calculate(stat = "slope")
```

```{r}
obs_slope
```

## Now

```{r obs-t}
#| echo: true

evals_lm <- lm(score ~ bty_avg,
               data = evals)

get_regression_table(evals_lm)
```
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
How did R calculate the $t$-statistic?
:::

::: panel-tabset
## Step 1: SE

::: columns
::: {.column width="40%"}
$SE_{b_1} = \frac{\frac{s_y}{s_x} \cdot \sqrt{1 - r^2}}{\sqrt{n - 2}}$
:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
```{r}
se_num <- ( sd(evals$score) / sd(evals$bty_avg) ) * 
  sqrt(
    1 - cor(evals$score, evals$bty_avg)
    ) 

se_denom <- sqrt(nrow(evals) - 2)

se <- se_num / se_denom

se
```
:::
:::

## Step 2: t-statistic

::: columns
::: {.column width="40%"}
$t = \frac{b_1}{SE_{b_1}}$
:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
```{r}
t <- pull(obs_slope) / se

t
```
:::
:::

## Proof!

```{r}
get_regression_table(evals_lm)
```
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
How does R calculate the p-value?
:::

. . .

```{r}
#| fig-align: center
#| out-width: 50%

ggplot(data = tibble(x = c(-5, 5)), 
       mapping = aes(x)) +
  stat_function(fun = dt, 
                args = list(df = nrow(evals) - 2), 
                color = "blue", 
                linewidth = 1.5) +
  ylab("") +
  scale_y_continuous(breaks = NULL) +
  geom_vline(xintercept = t, color = "red", linewidth = 1.5)

```

. . .

**How many degrees of freedom does this** $t$-distribution have?

## 

::: {style="font-size: 3.5em; color: #B6CADA;"}
Did we get similar results between these methods?
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
Why not always use theoretical methods?
:::

. . .

::: columns
::: {.column width="45%"}
Theory-based methods only hold if the sampling distribution is normally shaped.
:::

::: {.column width="5%"}
:::

::: {.column width="45%"}
The normality of a sampling distribution depends **heavily** on model conditions.
:::
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
What are these "conditions"?
:::

. . .

::: columns
::: {.column width="40%"}
For linear regression we are assuming...
:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
**L**inear relationship between $x$ and $y$

</br>

**I**ndepdent observations

</br>

**N**ormality of residuals

</br>

**E**qual variance of residuals
:::
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
**L**inear relationship between $x$ and $y$
:::

```{r}
#| fig-align: center

and_vertebrates %>%
  filter(species != "Cascade torrent salamander") %>% 
  ggplot(mapping = aes(x = length_1_mm, 
                       y = weight_g, 
                       color = species)) +
  geom_point() + 
  labs(x = "Length (mm)", 
       y = "Weight (g)") +
  my_theme
```

. . .

**What should we do?**

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
Variable transformation!
:::

```{r transform}
#| fig-align: center

and_vertebrates %>%
  filter(species != "Cascade torrent salamander") %>% 
  ggplot(mapping = aes(x = length_1_mm, 
                       y = weight_g, 
                       color = species)) +
  geom_point() + 
  labs(x = "Log Transformed Length (mm)", 
       y = "Log Transformed Weight (g)") +
  scale_x_log10() +
  scale_y_log10() +
  my_theme
```

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
**I**ndependence of observations
:::

::: columns
::: {.column width="40%"}
The `evals` dataset contains `r nrow(evals)` observations on `r distinct(evals, prof_ID) %>% nrow()` professors. Meaning, professors have **multiple** observations.

</br>

*What can we do?*
:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
**Best** -- use a random effects model

**Reasonable** -- collapse the multiple scores into a single score
:::
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
**N**ormality of residuals
:::

```{r non-normal}

fish <- and_vertebrates %>%
  filter(!species %in% c("Cascade torrent salamander", "Coastal giant salamander")) 

curved_lm <- lm(weight_g ~ length_1_mm, data = fish)

get_regression_points(curved_lm) %>% 
  ggplot(mapping = aes(x = residual)) + 
  geom_histogram() + 
  labs(x = "Residual") +
  my_theme
```

. . .

**What should we do?**

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
Variable transformation!
:::

```{r normal}

fish <- and_vertebrates %>%
  filter(!species %in% c("Cascade torrent salamander", "Coastal giant salamander"))

not_curved_lm <- lm(log(weight_g) ~ log(length_1_mm), data = fish)

broom::augment(not_curved_lm) %>% 
  mutate(.resid = `log(weight_g)` - `.fitted`) %>% 
  ggplot(mapping = aes(x = .resid)) + 
  geom_histogram(binwidth = 0.1) + 
  labs(x = "Residual") +
  xlim(c(-0.5, 0.5)) +
  my_theme
```

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
**E**qual variance of residuals
:::

```{r non-constant-variance}
#| fig-align: center

ggplot(data = hbr_maples, 
       mapping = aes(x = stem_length, y = stem_dry_mass)
       ) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    x = "Stem Length (mm)",
    y = "Stem Dry Mass (g)",
    title = "Sugar Maple Seedlings in Hubbard Brook LTER"
  ) +
  theme_minimal() + 
  my_theme
```

. . .

**What should we do?**

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
Variable transformation!
:::

```{r}
ggplot(data = hbr_maples, 
       mapping = aes(x = stem_length, y = stem_dry_mass)
       ) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    x = "Stem Length (mm)",
    y = "Log Transformed Stem Dry Mass (g)",
    title = "Sugar Maple Seedlings in Hubbard Brook LTER"
  ) +
  theme_minimal() + 
  scale_y_log10() +
  my_theme
```

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
Are these conditions required for **both** methods?
:::

. . .

::: columns
::: {.column width="40%"}
::: {style="font-size: 1.5em; color: #B6CADA;"}
Simulation-based Methods
:::

::: {style="font-size: 0.75em; color: #FFFFFF;"}
-   Linearity of Relationship

-   Independence of Observations

-   Equal Variance of Residuals
:::
:::

::: {.column width="5%"}
:::

::: {.column width="45%"}
::: {style="font-size: 1.5em; color: #B6CADA;"}
Theory-based Methods
:::

::: {style="font-size: 0.75em; color: #FFFFFF;"}
-   Linearity of Relationship
-   Independence of Observations
:::

::: {style="font-size: 0.75em; color: #B6CADA;"}
-   Normality of Residuals
:::

::: {style="font-size: 0.75em; color: #FFFFFF;"}
-   Equal Variance of Residuals
:::
:::
:::

## 

::: {style="font-size: 2em; color: #FFFFFF;"}
What happens if the conditions are violated?
:::

. . .

In general, when the conditions associated with these methods are violated, the permutation and $t$-distributions will underestimate the true standard error of the sampling distribution.
