---
title: "Philosophy of Statistics"
date: "September 23, 2021"
author: "Dr. Allison Theobold"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---        

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# R options
options(
  htmltools.dir.version = FALSE,
  tibble.width = 65,
  width = 65
  )

# figure height, width, dpi
knitr::opts_chunk$set(fig.width = 8, 
                      fig.asp = 0.618,
                      out.width = "60%",
                      dpi = 300)

# fontawesome
htmltools::tagList(rmarkdown::html_dependency_font_awesome())

# magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}

# xaringanExtra
library(xaringanExtra)
xaringanExtra::use_panelset()

library(emo)
library(palmerpenguins)
library(tidyverse)
```

```{r set-theme, include = FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color      = "#b76352", # mango
  secondary_color    = "#34605f", # bayberry
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  header_color = "#793540", #rhubarb
  white_color = "#F5F5F5", # lightest color
  black_color = "#36454F", # darkest color
  text_font_size = "30px", 
  link_color = "#696969" #grey
)
```

class: middle

.huge[Warm-up]

Sketch out what you recall the scientific method / process looking like.

---

class: inverse, center

.huge[.hand[Where does statistics happen?]]

---

class: inverse, center

.honey[
.huge[
.hand[
Why?
]
]
]

---

class: middle 

.bayberry[
.larger[For a scientific study, what is the purpose of a statistical analysis?] 
]

---

class: middle, inverse

.larger[What does it mean for an analysis to be objective?]

---

class: middle

.rhubarb[
.larger[What would help you feel more confident in someone else's statistical 
analysis?]
]

---

```{r, echo = FALSE, warning = FALSE, fig.align='center'}
penguins %>% 
  ggplot(aes(x = species, y = bill_length_mm)) + 
  geom_jitter() +
  labs(x = "Penguin Species", 
       y = "Bill Length (mm)")
```

