---
title: "Statistical Critique 2: Exploring p-values"
---

```{r}
#| label: data-load

# This is where you will load in your data! 
# Reference the R package you data live inside, since you will need to load that package in!

library(tidyverse)
library(moderndive)

# Package for your data:

```

# Part One: Revisiting the Midterm Project 

For the first part of this critique, you are going to revisit the model you selected for your Midterm Project. 

## Visualizations

**In the code chunk below, copy-and-paste the plots you created in your Midterm Project.**

```{r}
#| label: project-plots


```

## Model Justification

**Justify (in 2-3 sentences) why you chose this model!**

# Part Two: Using a p-value Instead

```{r}
#| label: obtain-p-values
# Step 1: fit the most complex regression model
# Step 2: run an ANOVA to obtain p-values

```

**Based on the p-value(s) you obtained, what model is the "best" model?**


# Part Three: Learning More about the Backlash Against $p$-values

**Type out what you find about your journal's publication criteria or simply copy-and-paste the criteria you find listed on their website.**


## Part Four: Lessons Learned


