---
title: "Week 9 -- Comparing Many Means (ANOVA)"
format: docx 
editor: visual
---

```{r}
#| include: false

library(tidyverse)
```

# Chapter 22

**What does ANOVA stand for?**

</br> </br>

**What is the null hypothesis for an ANOVA?**

</br> </br>

**What is the alternative hypothesis for an ANOVA?**

</br> </br>

**What are the conditions for an ANOVA?**

</br> </br>

## Section 2 -- Randomization Test for Comparing Many Means

### Conditions

**What are ways to assess the equal variability condition?**

</br> </br>

**If it appears that the variability between groups is similar can we say that the equal variance condition is met?**

</br> </br>

### Comparing every mean

**Why should we compare every mean rather than only the means that look the most different?**

</br> </br>

## Test statistic

**What is the statistic used in an ANOVA?**

</br> </br>

**How is this statistic calculated?**

</br> </br>

## Creating a permuted sample

Last week, we used cards to simulate what could have happened if $H_0$ was true by ripping the cards in half to separate the $x$ and $y$ values. We then shuffled the cards and created new ($x$, $y$) pairs by resampling from the data without replacement.

**How do we simulate what could have happened if** $H_0$ was true for an ANOVA?

</br> </br>

**What tail do we use to calculate the p-value for an ANOVA?**

</br> </br>

**If you reject** $H_0$ in an ANOVA what can you conclude?

</br> </br>

## Section 3 -- Mathematical Model

**What distribution do we use to calculate the p-value for a theory-based ANOVA?**

</br> </br>

```{r}
#| echo: false
lm(score ~ gender, data = moderndive::evals) %>% 
  anova() %>% 
  broom::tidy() %>% 
  knitr::kable()
```

**In the ANOVA table above, what does the `statistic` column represent?**

</br> </br>

**How was that `statistic` calculated? Specifically, what values from the table were used to obtain it?**

</br> </br>

**What distribution was used to calculate the `p.value`?**

</br> </br>
