---
title: "Lab 3: Grading Guide"
format: 
  html: 
    self-contained: true
editor: visual
execute: 
  eval: false
---

Q1\
Q2\
Q3\
Q4\
Q5\
Q6\
Q7\
Q8\
Q9\
Q10\
Q11\
Q12\
Q13

# Question 1 -- Size of `and_vertebrates` dataset

**Success**: 

- has `glimpse()` somewhere in their code
- states there are **32,209** rows and **16** columns

**Growing**: 

- If no code is present
- If they flip the rows and columns
- If they do not provide size of data

*Feedback for no code:* For Question 1, you needed to write code to explore the size of the data and the types of variables. 

*Feedback for no / incorrect dimensions:* For Question 1, look again at the output of the glimpse() function, how many rows and columns are included in the nycflights dataset?

# Question 2 -- Categorical variables in dataset

**Success**: 

- Lists all variables with a `chr` data type: 
  * `sitecode`
  * `section`
  * `reach`
  * `unittype`
  * `species`
  * `clip`
  * `notes`
- Permitted to miss 1 variable
  
**Growing**:

- If misses 2 or more variables
- If includes date variable (`sampledate`)

*Feedback:* For Question 2, you need to be careful to include ALL of the
variables that R believes are categorical. What data types did we say in class
that categorical variables can have?

# Question 3 -- Levels of the species variable

**Success**

- Provides code that uses `distinct()` function
- States there are 4 levels: Cutthroat trout, Coastal giant salamander, Olympic
torrent salamander, and NA

**Growing:** Does not use `distinct()` function

*Feedback:* For Question 3, there is a function which finds the *distinct* levels
of a categorical variable. What function might that be?

# Question 4 -- Filtering to including only trout

**Success:** Code should look similar to:

```{r fish-filter-1}
trout <- filter(and_vertebrates, 
                species == "Cutthroat trout")
```

::: {.callout-note}
The can use %>% or pass in `and_vertebrates` as first argument
:::

**Growing:** If `filter()` is not correct

# Question 5 -- Filtering to include only catchable trout

**Success:** Code should look similar to:

```{r fish-filter-2}
trout <- filter(trout, 
                length_1_mm > 101)
```

**Growing:** If `filter()` is not correct 

- using `and_vertebrates` 

- using the wrong variable / value 

- making a new dataset instead of updating the `trout` dataset

*Feedback:* Careful! In Question 5 you were given code that would update the 
trout dataset, *not* make a new dataset. By making a new dataset, you forgot to 
use the correct name of the dataset later on! 

## Question 6 -- Distribution of trout lengths 

**Success**: Acceptable `geom`s:

  * `geom_histogram()`
  * `geom_dotplot()` (though not great)
  * `geom_density()` 

::: {.callout-note}
Note: Axis labels are not required! But, if they don't change it, give the 
following feedback:

*For Question 6, it is always nice to have professional looking axis labels!*
:::

**Growing**: 

- Uses `geom_boxplot()`

*Feedback:* For Question 6, we need to use a geom which allows for us to see the *shape* of the distribution. Boxplots hide distributions with multiple modes, so what type of plot would be better?


## Question 7 -- Sources of variation

**Success:** Names three "reasonable" sources of variation in fish length

**Growing:** Names an unreasonable source of variation in fish lengths

*Feedback:* For Question 7, we are interested in variables that, if changed, we
would expect changes in the length of the Cutthroat trout. 

## Question 8 -- Ridge plot

**Success:** Code should look like the following

```{r}
ggplot(data = fish, 
       mapping = aes(x = length_1_mm, y = unittype)) +
  geom_density_ridges() +
  labs(x = "Length (mm)", 
       y = "Channel Section")
```

**Growing:** 

- Doesn't change axis labels

*Feedback:* For Question 8, you were asked to give your visualization nice axis
labels. Note that means BOTH axes!  

- Doesn't include units (mm) in x-axis label

*Feedback:* For Question 8, it is important for the axis label to contain the
units of the variable. What units were the lengths measured in?

- Y-axis should say something about the type of channel

*Feedback:* For Question 8, it is important for the axis label to describe the variable that is being measured.

## Question 9 -- Adding another categorical variable

**Success:** Uses either color or facets to incorporate `species`

```{r channel-section}
## Option 1
ggplot(data = fish, 
       mapping = aes(x = length_1_mm, y = unittype)) +
  geom_density_ridges() +
  facet_wrap(~species)

## Option 2
ggplot(data = fish, 
       mapping = aes(x = length_1_mm, y = unittype, color = species)) +
  geom_density_ridges() 

```

**Growing:** If doesn't use facets or colors

::: {.callout-note}
If they got a growing on #8 for an axis label, they don't get a growing here. 
:::

**Question 10** -- Based on your plot, how different are the lengths of the Cutthroat trout between the different channel types and forest sections?

## Question 10 -- Based on the plot, how different are the lengths between the channel types and forest sections?

**Success:** Must have all of the following

- Comparisons between forest sections (centers, spreads, shapes)
- Comparisons between channel types
- State what channel types did not have both types of forest

**Growing:**

- Comparisons are incomplete
- Does not state what channel types did not have both types of forest

*Feedback:* Careful! For Question 10 you should note if not every forest type 
was observed for every type of channel.

## Question 11 -- Average length of Cutthroat trout between channel types

**Success:** Code should look similar to:

```{r mean-length}
trout %>%
  group_by(unittype) %>%
  summarize(mean_length = mean(length_1_mm))
```

If they don't name their summary statistics

*Feedback:* For Question 11, the output of your summary statistics looks a lot
nicer if you give them names! (like the example that was given)

**Growing:** If process is not correct

## Question 12 -- Find the average length of Cutthroat trout between channel
types and forest section

**Success:** Code should look similar to:

```{r mean-length-2}
trout %>%
  group_by(unittype, species) %>%
  summarize(mean_length = mean(length_1_mm))
```

**Growing:** If process is not correct

## Question 13 -- Differences in averages compared to plot

**Success:**

- States that the averages are all fairly similar
- 

**Growing:** If all they say is that the differences in means are similar to the
visualization 

*Feedback:* For Question 13, *how* are the differences similar to the
visualizations? How does what you saw in the visualizations manifest in these
means? 


