---
title: "Lab 4: Grading Guide"
format: 
  html: 
    self-contained: true
editor: visual
execute: 
  eval: false
---

Q1 - \
Q2 - \
Q3 - \
Q4 - \
Q5 - \
Q6 - \
Q7 - \
Q8 - \
Q9 - \
Q10 - \


# Question 1 -- Size of `ntl_icecover` dataset

**Success**:

-   has `glimpse()` somewhere in their code
-   states there are **334** rows and **5** columns

**Growing**:

-   If no code is present
-   If they provide incorrect size of data

::: {.callout-note}
If they say there are 431 rows, they got their answer from the help file! 

*Feedback:* Careful! The best way to obtain information about the dataset is using the glimpse() function! 
:::

# Question 2 -- Scatterplot

**Success:** Code should look like the following

```{r}
ggplot(data = ntl_icecover, 
       mapping = aes(y = ice_duration, x = year)) +
  geom_point() +
  labs(x = "Year", 
       y = "Ice Duration (days)")
```

**Growing:**

-   Doesn't change axis labels

*Feedback:* For Question 2, you were asked to give your visualization nice axis labels. Note that means BOTH axes!

-   Doesn't include units (days) in y-axis label

*Feedback:* For Question 2, it is important for the axis label to contain the units of the variable. What unit was ice duration measured in?


# Question 3 -- Describe the regression line

**Success**: Addresses form, direction, strength and unusual points.

- Form: Linear
- Direction: Negative
- Strength: Moderate
- Outliers: Maybe the 2001 observation with ~20 days of ice?


**Growing**:

- If they don't discuss one of the for aspects.

*Feedback:* For Question 3, you were asked to describe the form, direction, strength, and unusual points. Make sure you discuss all four aspects!

- If they say there are unusual points / outliers but don't say where.

*Feedback:* For Question 3, if you believe there are unusual observations you need to tell the reader where these points are! 


# Question 4 -- Add a regression line

**Success**:

-   Includes regression line with `geom_smooth()` **and** `method = "lm"` option

::: {.callout-note}
They don't need to turn the standard errors off (`se = FALSE`), but great if they did! 
:::

**Growing**:

- If they do not add regression line

*Feedback:* For Question 4, you needed to add a regression line to the previous scatterplot!

- If they do not use `method = "lm"` to get straight line

*Feedback:* For Question 4, you need to have a **straight** line, not a wiggly line! 

# Question 5 -- Find correlation

**Success:** Code should look like the following

```{r}
get_correlation(data = ntl_icecover, ice_duration ~ year, na.rm = TRUE)
```

**Growing:**

- Doesn't remove NAs

*Feedback:* For Question 5, it is important to remove the NAs before calculation the correlation! Is there another argument (input) for the get_correlation() function that allows you to remove the missing values?

# Question 6 -- Fit linear regression

**Success:** Code should look like the following

```{r}
my_model <- lm(ice_duration ~ year, data = ntl_icecover)
```

**Growing**:

- Swaps response and explanatory

*Feedback:* Careful! With the `lm()` function the response variable goes first (before the `~`) and the explanatory comes second (after the `~`).

# Question 7 -- Get regression table

**Success:** Code should look like the following

```{r}
get_regression_table(my_model)
```

**Growing**:

- Doesn't use the `get_regression_table()` function

*Feedback:* For Question 7, we want out model output to look tidy, which is why we use the `get_regression_table()` function! 

# Question 8 -- Write out regression equation

**Success**:

$$ \widehat{\text{ice duration}} = 495 - 0.203 \times (\text{year})$$

- Indicates the response (ice duration not y) is estimated / predicted / the mean with either a hat or with language
- Inputs values in correct location
- References `year`(not x)

**Growing**:

- Does not indicate the response (ice duration) is estimated / predicted / the mean

*Feedback:* For Question 8, remember the estimated regression equation has a hat over y. What does that hat mean? How can you incorporate that into your regression equation? 

- Uses x and y instead of variable names

*Feedback:* When writing out the regression equation, you need to reference the context of the line that was found. What variable was the response? What variable was the explanatory? 

::: {.callout-note}
If they use x and y in the equation but then define what variables they are associated with, that is okay!
:::


# Question 9 -- Interpret slope 

**Success**:

- Correct interpretation of slope for 1 year increase
  * 1 year increase
  * mean / predicted / estimated ice duration 
  * decreases by 0.203 days

- Correct interpretation of slope for 100 year increase
  * 100 year increase
  * mean / predicted / estimated ice duration 
  * decreases by 20.3 days
  
**Growing**:

- Incorrect interpretation of the slope

*Feedback:* For slope interpretations, we always increase the explanatory variable by 1 unit and the slope coefficient is the expected change in the mean of the response variable.

- Does not interpret the slope for a 100 year change.

*Feedback:* For Question 9, you were asked to interpret the slope for a 100 year increase, too!

# Question 10 -- Multivariate plot

**Success:** Code should look like the following

```{r}
ggplot(data = ntl_icecover, 
       mapping = aes(y = ice_duration, x = year, color = lakeid)) +
  geom_point() +
  labs(x = "Year", 
       y = "Ice Duration (days)")
```

**Growing**:

????