---
title: "Lab 9 -- Grading Guide"
format: 
  html:
    self-contained: true
editor: visual
execute: 
  eval: false
---

Q1:\
Q2:\
Q3:\
Q4:\
Q5:\
Q6:\
Q7:\
Q8:\
Q9:\
Q10:\
Q11:\
Q12:\
Q13:\
Q14:\
Q15:\
Q16:\
Q17:\
Q18:\
Q19:\
Q20:\

## Question 1

**To earn a Success:**

- fills in code
- adds axis label

::: {.callout-warning}
Their axis label needs to indicate this is a score for Grade 8 Math
:::

*If they do not have "grade 8" or "math score" in their title:*

> What context are these math scores for? What exam? What grade?

## Question 2

**To earn a Success:**

- compares centers of plots
- compares spreads of plots


## Question 3

**To earn a Success:**

- fills in code
- adds axis label

::: {.callout-warning}
Their axis label needs to indicate this is a score for Grade 8 Math
:::

*If they do not have "grade 8" or "math score" in their title:*

> What context are these math scores for? What exam? What grade?

## Question 4 

**To earn a Success:**

- compares centers of plots
- compares spreads of plots

## Question 5 -- Important!

**To earn a Success:**

::: columns
::: {.column width="40%"}
Using fill:

```{r}
ggplot(data = math_scores, 
       mapping = aes(x = grade_8_math_score, 
                     y = year_cat, 
                     fill = continent)) +
  geom_density_ridges(alpha = 0.5, scale = 1) + 
  labs(x = "Grade 8 Math Score on TIMSS", 
       y = "Year", 
       fill = "Continent")
```

:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
Using facets:

```{r}
ggplot(data = math_scores, 
       mapping = aes(x = grade_8_math_score, 
                     y = year_cat)) +
  geom_density_ridges(alpha = 0.5, scale = 1) + 
  labs(x = "Grade 8 Math Score on TIMSS", 
       y = "Year") + 
  facet_wrap(~continent)
```

:::
:::

::: {.callout-warning}
Their axis label needs to indicate this is a score for Grade 8 Math
:::

*If they don't change the transparency of their plots:* 

> Remember you can use alpha to change the transparency of the density ridges so you can see them all!

*If they do not have "grade 8" or "math score" in their title:*

> What context are these math scores for? What exam? What grade?

## Question 6 -- Important! 

**To earn a Success:** compares if distributions of continents move over time or if they stay similar 

## Question 7 -- Important!

**To earn a Success:** 

- compares observations within a continent
- notices countries are recorded multiple times
- states the condition is violated

*If they do not state the condition is violated:*

> Look at the observations for Africa! How many observations for Ghana are there? Is it reasonable to assume those observations are independent? 

*If they say observations for continents are related in time:*

> Do we have one observation per continent? If not, what are the observations of? Are these observations within a continent related? 

## Question 8 -- Important!

**To earn a Success:**

- connects comparison with observations within a year or between continents

::: columns
::: {.column width="40%"}
- says countries / continents are not related 
- justifies decision (no)
:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
- says countries / continents are related 
- justifies decision (spatial relationship)
:::
:::

*If they do not justify their decision:*

> An essential part of evaluating model conditions is to justify the decisions you make. What was your reasoning for your decision?

## Question 9 -- Important!

**To earn a Success:**

- compares observations between continents
- states that a country can only belong to one continent
- says condition is not violated

*If they don't discuss how many continents a county can belong to:*

> How many continents can a country belong to? What does that say about between continent independence?

*If they do not justify their decision:*

> An essential part of evaluating model conditions is to justify the decisions you make. What was your reasoning for your decision?

## Question 10 -- Important!

**To earn a Success:**

- compares observations between years
- states that observations are related in time
- says condition is violated

*If they do not state the condition is violated:*

> Is the observation for Ghana in 2003 related to the observation for Ghana in 2007?

*If they do not justify their decision:*

> An essential part of evaluating model conditions is to justify the decisions you make. What was your reasoning for your decision?

## Question 11 -- Important!

**To earn a Success:**

- discusses the shape of the distributions 
- says some distributions have more than one mode
- says condition is violated

*If they do not state the condition is violated:*

> How many peaks does the Normal distribution have? How many peaks do you see in some of these distributions?

*If they don't make a decision (violated / not):* 

> A critical part of evaluation model conditions is to make a decision! Do you feel that this condition is or is not violated?

## Question 12 -- Important!

**To earn a Success:**

- compares the log variances between continents / years
- says there are large differences in the log variances
- says condition is violated

*If they do not state the condition is violated:*

> How many times larger is the largest variance than the smallest variance? If it is less than 2, they could be similar values, but larger than 2 they start to seem pretty different. 

## Question 13 

**To earn a Success:** Code should look like:

```{r}
aov(grade_8_math_score ~ year, data = math_scores) %>% 
  broom::tidy()
```

*If they swap $y$ and $x$:*

> Careful! In the lm() function the response goes first! 


## Question 14

**To earn a Success:**

- reject $H_0$
- state that the p-value was less than 0.1

::: {.callout-note}
# Fine to not include p-value

*If they don't state what the p-value was:*

> Remember it is important to be transparent with your decision, what p-value did you obtain for this test?
:::

*If they don't state their $\alpha$:*

> What alpha was used for your decision? 

## Question 15 

**To earn a Success:** Conclude at least on year has a different mean grade 8 math score

*If they say all the years have different means:*

> What is the alternative hypothesis for a one-way ANOVA? Is it that all of the years have different means?

## Question 16 

**To earn a Success:** Code should look like:

```{r}
null_dist <- math_scores %>%
  specify(response = grade_8_math_score, explanatory = continent) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "F")
```

## Question 17 

**To earn a Success:** Code should look like:

```{r}
visualize(null_dist) +
  shade_p_value(obs_stat = obs_F, direction = "greater")
```

*If they don't use `"greater"` in their direction:*

> Careful! F-statistics can only be positive! What direction should we use to calculate the p-value for statistics that cannot be negative?

## Question 18

**To earn a Success:** Code should look like:

```{r}
get_p_value(null_dist, 
            obs_stat = obs_F, 
            direction = "greater")
```

*If they don't use `"greater"` in their direction:*

> Careful! F-statistics can only be positive! What direction should we use to calculate the p-value for statistics that cannot be negative?

## Question 19

**To earn a Success:**

- reject $H_0$
- state that the p-value was less than 0.1

::: {.callout-note}
# Fine to not include p-value

*If they don't state what the p-value was:*

> Remember it is important to be transparent with your decision, what p-value did you obtain for this test?
:::

*If they don't state their $\alpha$:*

> What alpha was used for your decision? 

## Question 20

**To earn a Success:** Conclude at least on continent has a different mean grade 8 math score

*If they say all the continents have different means:*

> What is the alternative hypothesis for a one-way ANOVA? Is it that all of the continents have different means?

