---
title: "Lab 8 Grading Guide"
format: html
editor: visual
---

## Question 1

**To earn a Success**: Creates a scatterplot with the following qualities:

-   life expectancy on y-axis
-   GDP per capita on the x-axis
-   axis labels for both axes, including units

**Growing:**

*If they swap* $x$ and $y$:

> Careful! What variable should be the response variable (located on the y-axis)?

*If they don't include axis labels*:

> Careful! Your axis labels should also include the unit each variable was measured in! If you are unsure what unit a variable was measured in, you can consult the data documentation using ?gapminder. :)

## Question 2

**To earn a Success**:

-   takes log of GDP (using `scale_x_log10()`)
-   renames x-axis label to indicate log GDP (or log \$) is what is being plotted

**Growing:**

*If they log transform life expectancy (y)*:

> Similar to the Principle of Parsimony, we want to transform as few variables as is necessary. Look back at the plot with just x log transformed. Is log transforming y and x notably better?

*If their x-axis doesn't indicate the log was taken*:

> Careful! You axis label needs to indicate how the original variable was transformed!

## Question 3

**To earn a Success**: Code should look like the following:

```{r}
#| eval: false
#| echo: true
gapminder_lm <- lm(lifeExp ~ log(gdpPercap), data = gapminder)
```

**Growing:**

*If they swap `lifeExp` and `gdpPercap`*:

> Careful! What variable comes first in the lm() function? The response or explanatory variable?

*If they don't use `log(gdpPercap)` when fitting their regression*:

> What variable did you decide to transform in #2? How should that variable appear in the regression model you are fitting here?

## Question 4

**To earn a Success**:

-   Says it **is not** reasonable to assume the observations are independent

-   Describes how observations are not independent citing **at least one** of the following:

    -   each country has repeated observations
    -   observations for each country are related in time (temporal correlation)
    -   countries of close geographical proximity may share information (spatial correlation)

**Growing:**

*If their justification doesn't talk about the context of the data (e.g., temporal relationships between observations)*:

> Your justification needs to make direct reference to the context of the data.

*If their reasoning doesn't include any of the above justifications*:

> How many observations are their for each country? Are these observations related in some way? If so, how?

## Question 5

**To earn a Success**:

::: columns
::: {.column width="45%"}
-   Says the condition is violated
-   Justifies with the left skew of the distribution
:::

::: {.column width="5%"}
:::

::: {.column width="45%"}
-   Says the condition is not violated
-   Justifies with characteristics of the distribution
:::
:::

**Growing:**

*If their justification is insufficient*:

> When evaluating conditions the choices are subjective, so it is necessary for you to justify WHY you made the decision you did. Your justification should make direct reference to characteristics of the distribution of residuals.

## Question 6

**To earn a Success**:

-   Says the condition is violated
-   References how the residuals change / decrease for larger values of log(GDP)

**Growing:**

*If they say the condition is not violated*:

> Equal variance requires that the spread of residuals / the vertical width (e.g., going from -20 to +20) stays the same for ALL values of the explanatory variable (across the x-axis). Is that the case? Why or why not?

*If they say the condition is / is not violated but reference the number of observations above and below the line*:

> Equal variance is not about having equal spread of points above and below the line -- it is okay for there to be more residuals below the line compared to above the line. They key is that the spread of residuals / the vertical width (e.g., going from -20 to +20) stays the same for ALL values of the explanatory variable (across the x-axis). Is that the case? Why or why not?

*If they say the condition is violated but have insufficient justification as to why*:

> When evaluating conditions the choices are subjective, so it is necessary for you to justify WHY you made the decision you did. Your justification should make direct reference to characteristics of the plot of the residuals versus fitted values. Specifically, we are evaluating if the spread of residuals / the vertical width (e.g., going from -20 to +20) stays the same for ALL values of the explanatory variable (across the x-axis). Is that the case? Why or why not?

## Question 7

**To earn a Success**:

-   $H_0$: there is no linear relationship between log GDP per capita and life expectancy
-   $H_A$: there is a linear relationship between log GDP per capita and life expectancy

**Growing:**

*If they say GDP instead of log GDP*:

> Careful! How did you transform your variable(s) in #2? What variables are you looking at the linear relationship between?

*If they say there is a **positive** relationship in their alternative*:

> The standard hypothesis test for the slope uses a two-sided alternative hypothesis, unless we knew 100% going in that the relationship between $x$ and $y$ was positive. Did you know that the relationship was positive BEFORE you made your visualizations?

## Question 8

**To earn a Success**: Code should look like the following:

```{r}
#| echo: true
#| eval: false
obs_slope <-  gapminder %>%
  specify(response = lifeExp, explanatory = log(gdpPercap)) %>%
  calculate(stat = "slope")
```

**Growing:**

*If they swap `lifeExp` and `gdpPercap`*:

> Careful! What variable is your response variable?

*If they use `GDP` instead of `log(GDP)` for their explanatory:*

> Careful! You need to be consistent with the transformation you decided in #2.

## Question 9

**To earn a Success**: Code should look like the following:

```{r}
#| echo: true
#| eval: false
null_dist <- gapminder %>%
  specify(response = lifeExp, explanatory = log(gdpPercap)) %>%
  hypothesise(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "slope")
```

**Growing:**

*If they also don't use the log in this step:*

> Update your code to use the same variable transformation that you use in #8.

## Question 10

**To earn a Success**: Code should look like the following:

```{r}
#| eval: false
#| echo: true
get_p_value(null_dist, 
            obs_stat = obs_slope, 
            direction = "two-sided")
```

:::{.callout-note}
# Consistent with alternative hypothesis

If they said the relationship was positive in the alternative, then their `direction` should be `"greater"`. 
:::

**Growing:**

*If they don't use a direction that is consistent with what they said in their alternative (`"two-sided"` or `"greater"`):

> How many tails are there in the hypotheses stated in #7?

## Question 11

**To earn a Success**: States they reject $H_0$ because the p-value is less than 0.1

**Growing:**

*If they don't reference their significance level:*

> Careful! Hypothesis test decisions can differ based on the significance threshold that was used. What threshold did you use?

*If they do not make a decision*:

> You were asked to make a decision regarding the hypotheses, which has two possible options. Which option do you choose and why?

## Question 12

**To earn a Success**: Code should look like the following:

```{r}
#| eval: false
#| echo: true
get_regression_table(gapminder_lm, conf.level = 0.9)
```

::: callout-tip
# It's okay if their `conf.level` isn't 0.9

This doesn't need to be consistent, since we aren't making a confidence interval!
:::

## Question 13

**To earn a Success**: States that the p-value is essentially the same

## Question 14

**To earn a Success**:

-   states that the permutation distribution is approximately Normal
-   states that a t-distribution is a reasonable approximation

**Growing:**

*If they don't reference the permutation distribution*:

> Theory-based methods are an approximation for simulation-based methods. Based on the permutation distribution you created, do you believe a t-distribution is a good approximation for THIS permutation distribution? Why or why not?

## Question 15

**To earn a Success**: Answer must agree with what they said in #4-6

::: columns
::: {.column width="30%"}
If they said equal variance and / or independence was violated, they must say that **neither** p-value is reliable.
:::

::: {.column width="2%"}
:::

::: {.column width="30%"}
If they said normality **was** violated but equal variance and independence **were not** violated, they must say that the **simulation-based** p-value is more reliable.
:::

::: {.column width="2%"}
:::

::: {.column width="30%"}
If they said none of the conditions were violated, they must say that **both** p-values are reliable.
:::
:::

**Growing:**

*If they choose the wrong method*:

> Look back at the conditions required for each of these methods. Which conditions did you say were violated? What does that imply for the method(s) which give you the most reliable p-value?
