---
title: "Lab 4: Grading Guide"
format: html
editor: visual
execute: 
  eval: false
---

Q1 -\
Q2 -\
Q3 -\
Q4 -\
Q5 -\
Q6 -\
Q7 -\
Q8 -\
Q9 -\
Q10 -\
Q11 -\

# Question 1 -- Size of `ntl_icecover` dataset

**Success**:

-   has `glimpse()` somewhere in their code
-   states there are **334** rows and **5** columns

**Growing**:

-   If no code is present
-   If they provide incorrect size of data

::: callout-note
If they say there are 431 rows, they got their answer from the help file!

*Feedback:* Careful! The best way to obtain information about the dataset is using the glimpse() function!
:::

# Question 2 -- Scatterplot

**Success:** Code should look like the following

```{r}
ggplot(data = ntl_icecover, 
       mapping = aes(y = ice_duration, x = year)) +
  geom_point() +
  labs(x = "Year", 
       y = "Ice Duration (days)")
```

**Growing:**

-   Doesn't change y-axis labels

*Feedback:* For Question 2, you were asked to give your visualization nice axis labels. Also, it is important for the axis label to contain the units of the variable. What unit was ice duration measured in?

-   Doesn't include units (days) in y-axis label

*Feedback:* For Question 2, it is important for the axis label to contain the units of the variable. What unit was ice duration measured in?

# Question 3 -- Describe the regression line

**Success**: Addresses form, direction, strength and unusual points.

-   Form: Linear
-   Direction: Negative
-   Strength: Moderate
-   Outliers: Maybe the 2001 observation with \~20 days of ice?

**Growing**:

-   If they don't discuss one of the for aspects.

*Feedback:* For Question 3, you were asked to describe the form, direction, strength, and unusual points. Make sure you discuss all four aspects!

-   If they say there are unusual points / outliers but don't say where.

*Feedback:* For Question 3, if you believe there are unusual observations you need to tell the reader where these points are!

# Question 4 -- Add a regression line

**Success**:

-   Includes regression line with `geom_smooth()` **and** `method = "lm"` option

::: callout-note
They don't need to turn the standard errors off (`se = FALSE`), but great if they did!
:::

**Growing**:

-   If they do not add regression line

*Feedback:* For Question 4, you needed to add a regression line to the previous scatterplot!

-   If they do not use `method = "lm"` to get straight line

*Feedback:* For Question 4, you need to have a **straight** line, not a wiggly line! Take a look at the R resources to see how you specify a straight line!

# Question 5 -- Find correlation

**Success:** Code should look like the following

```{r}
get_correlation(data = ntl_icecover, ice_duration ~ year, na.rm = TRUE)
```

**Growing:**

-   Doesn't remove NAs

*Feedback:* For Question 5, it is important to remove the NAs before calculation the correlation! Is there another argument (input) for the `get_correlation()` function that allows you to remove the missing values? Hint: it looks like the option we used to remove missing values in the `mean()` function!

# Question 6 -- Fit linear regression

**Success:** Code should look like the following

```{r}
my_model <- lm(ice_duration ~ year, data = ntl_icecover)
```

**Growing**:

-   Swaps response and explanatory

*Feedback:* Careful! With the `lm()` function the response variable goes first (before the `~`) and the explanatory comes second (after the `~`).

# Question 7 -- Get regression table

**Success:** Code should look like the following

```{r}
get_regression_table(my_model)
```

**Growing**:

-   Doesn't use the `get_regression_table()` function

*Feedback:* For Question 7, we want out model output to look tidy, which is why we use the `get_regression_table()` function!

# Question 8 -- Write out regression equation

**Success**:

$$ \widehat{\text{ice duration}} = 495 - 0.203 \times (\text{year})$$

-   Indicates the response (ice duration not y) is estimated / predicted / the mean with either a hat or with language
-   Inputs values in correct location
-   References `year`(not x)

**Growing**:

-   Does not indicate the response (ice duration) is estimated / predicted / the mean

*Feedback:* For Question 8, remember the estimated regression equation has a hat over y. What does that hat mean? How can you incorporate that into your regression equation?

-   Uses x and y instead of variable names

*Feedback:* When writing out the regression equation, you need to reference the context of the line that was found. What variable was the response? What variable was the explanatory?

::: callout-note
If they use x and y in the equation but then define what variables they are associated with, that is okay!
:::

# Question 9 -- Interpret slope

**Success**:

-   Correct interpretation of slope for 1 year increase
    -   1 year increase
    -   mean / predicted / estimated ice duration
    -   decreases by 0.203 days
-   Incorrect interpretation of the slope

*Feedback:* For slope interpretations, we always increase the explanatory variable by 1 unit and the slope coefficient is the expected change in the mean of the response variable.

-   Doesn't interpret the value of -0.203

*Feedback:* Careful! You did not interpret what the value of -0.203 means in the context of the regression line!

-   States the duration of ice increases

*Feedback:* Look at the sign associated with the slope coefficient! What does the sign tell you about whether ice duration should increase or decrease?

-   Doesn't indicate the response is mean / estimate / predicted ice duration

*Feedback:* Careful! Your interpretation sounds like it is guaranteed that every 1 year increase in time will result in **exactly** a 0.203 day decrease in ice duration. Will there always be exactly this decrease? How do we indicate that our slope estimate is not exact?

# Question 10 -- 100 Year Increase

-   Correct interpretation of slope for 100 year increase
    -   100 year increase
    -   mean / predicted / estimated ice duration
    -   decreases by 20.3 days

**Growing**:

-   Uses y-intercept when increasing 100 years

*Feedback:* When we increase 100 years, technically the only aspect of the equation that changes is the amount of time our slope is multiplied by. The y-intercept does not change! With that in mind, how does a 100 year increase in time affect what you would expect for the duration of the ice on these lakes?

-   Don't provide an estimate of the change

*Feedback:* You need to use your estimated regression equation to state what the expected change in the ice duration would be if the number of years was increased by 100.

# Question 11 -- Multivariate plot

**Success:** Code should look like the following

```{r}
ggplot(data = ntl_icecover, 
       mapping = aes(y = ice_duration, x = year, color = lakeid)) +
  geom_point() +
  labs(x = "Year", 
       y = "Ice Duration (days)")
```

::: callout-tip
If they got a Growing for Question 4 from using a wiggly line, they don't need to get a Growing here!
:::

**Growing**:

????
