---
title: "Midterm Project Data Formatting"
---

Although R is a component of this course, it is not one of the learning outcomes. Thus, I have decided we **will not** be spending time writing the R code to wrangle your data into the format R expects them to be in. Instead, you are responsible for formatting your data so they are ready to be analyzed in class on Thursday. 

## Required Format

R expects every dataset to be in "tidy" format, meaning

- every column is a variable 
- every row is an observation
- every cell has one value

Each variable you selected to analyze for your midterm project will need to be 
included as **one** column in your dataset. 

For example, suppose the `ntl_icecover` data we worked with for Lab 4 had one row for each lake and column for every (displayed below): 

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: wide-data

library(tidyverse)
library(lterdatasampler)

ntl_icecover %>% 
  select(-ice_on, -ice_off) %>% 
  pivot_wider(names_from = year, values_from = ice_duration) %>% 
  select(1:12) %>% 
  gt::gt()

```

These data **could not** be analyzed by R, as the variable `year` occupies more than one column. To format these data, we would need to create **two** news columns, one to store the years and one to store the ice durations. The reformatted data would look like:

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: long-data

ntl_icecover %>% 
  select(-ice_on, -ice_off) %>% 
  pivot_wider(names_from = year, values_from = ice_duration) %>% 
  select(1:12) %>% 
  pivot_longer(-lakeid, names_to = "year", values_to = "ice_duration") %>% 
  gt::gt()

```


## How to Format Your Data

To get your data in the correct format, you have two options:

1. copy-and-paste each variable into a column in a new dataset
2. write the R (or Python) code to reformat your data

You should not spend time Googling or using Chat GPT to write the R code to format your data. I **will not** provide assistance writing R code to format your data. Yes, copying and pasting is a pain, but this is a lesson in good data storage practices. 

## File Format

You are expected to save your new dataset as either an Excel file (`.xlsx`) or 
as a comma separated values file (`.csv`). No other format will be accepted. 

## Example Video

In case you are unsure what I mean by "copy-and-paste" your variables into a dataset, I have recorded this short video walking you through how I would reformat a dataset to be ready for analysis. 

Link to Zoom recording: <>