---
title: "Welcome to RStudio Cloud!"
author: "Your Name Here!"
date: "January 12, 2023"
format: html
embed-resources: true
standalone: true
editor: visual
execute: 
  echo: true
---

## Quarto

This is a Quarto document.

Quarto is a software that allows you to interweave text and R code to create HTML, PDF, and Microsoft Word documents

The document may look a bit different that you are used to, but we will break it down!

Section headers are defined with hashtags (#), where one \# is the largest header, two \## is the second largest and so on.

You can make words boldface with two \*\* or two \_\_ (like **so** or **so**).

You can make words italicized with one \* or one \_ (like *so* or *so*).

You can make lists like so:

-   item 1
-   item 2
-   item 3

<!-- This is an comment. It WILL NOT show in your knitted document. -->

```{=html}
<!-- Comments are great ways to leave yourself notes inside your Quarto
document that no one can see!  -->
```
## R Code

You can differentiate the R code within a Quarto file from the body of the document, from the three tick marks (\`\`\`) they start and end with.

Here is an example of an R code chunk:

```{r packages}
#| include: false

library(tidyverse)
## Code to load in package we need
## <- this creates a code comment. R knows this IS NOT code because it has a # in front! 
```

Notice that after the \`\`\` you see a curly brace with an "r" in it ({r}), this indicates to Quarto that the code inside of that code chunk is R code (not Python code or another type of code).

Also, the "packages" after the "r" is the code chunk's name. It is good practice to name your code chunks based on the action they perform. It makes it easier to navigate to them in the navigation pane.

## Rendering

When you click the **Render** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

**Question 1:** Do you see the above code chunk when you knit the document? Why do you think this is the case?

## Including Code Output

You can include code output in your knitted document:

```{r cars}
glimpse(mpg)
```

**Question 2:** What do you think the above code does? What output does it give you?

## Including Plots

You can also embed plots in the knitted document.

Here is an example of a plot.

```{r pressure}
#| echo: false

ggplot(data = mpg, 
       mapping = aes(y = manufacturer, x = hwy)) + 
  geom_boxplot() +
  labs(x = "Highway Miles Per Gallon", 
       y = "Car Manufacturer")
```

**Question 3**: What do you think the `echo: false` option does in the above code chunk?

**Question 4:** What do you think the `mapping = aes(y = manufacturer, x = hwy))` code does?

**Question 5:** What do you think the `labs(x = "Highway Miles Per Gallon", y = "Car Manufacturer")` code does?
