---
title: "Model Justification"
format: html
editor: visual
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
#| include: false

library(tidyverse)
library(openintro)
library(moderndive)

```

This study seeks to investigate how the relationship between average SAT math scores and the percentage of students who are economically disadvantaged differs based on the size of the school. As seen in the plot below, the slope for this relationship was found to be similar across the different levels of schools. Therefore, this study used a parallel slopes (additive) multiple linear regression model.

```{r}
MA_schools %>% 
  ggplot(mapping = aes(x = perc_disadvan, y = average_sat_math, 
                       color = size)) + 
  geom_point(alpha = 0.5) + 
  geom_parallel_slopes(alpha = 0.5) + 
  labs(x = "Percentage of Students Economically Disadvantaged", 
       y = "Average SAT Math Score", 
       color = "Size of School")
```

**What do you feel is missing from this model justification?**
