---
title: ""
format: docx
editor: visual
---

```{r set-up}
library(palmerpenguins)
library(moderndive)
```

## Full Model -- ALL Explanatory Variables Included

```{r full}
lm(body_mass_g ~ . , data = penguins) %>% 
  get_regression_summaries()
```

::: {.callout-note}
# Starting adjusted $R^2$

Note the adjusted $R^2$ value you are starting with, as this will influence your decisions.
:::

## Variable Selection

Now, starting with our full model, we will use backwards selection to decide what variable(s) should be removed from the model. 

You can only delete a variable if it **increases** adjusted $R^2$. 

::: {.callout-note}
Note we are not saying how much adjusted $R^2$ needs to be increased, simply that it must be bigger. 
:::

## Models Deleting One Explanatory Variable

```{r minus-year}
lm(body_mass_g ~ . -year, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.872 -->

```{r minus-sex}
lm(body_mass_g ~ . -sex, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.849 -->

```{r minus-flipper-length}
lm(body_mass_g ~ . -flipper_length_mm, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.86 -->

```{r minus-bill-depth}
lm(body_mass_g ~ . -bill_depth_mm, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.87 -->

```{r minus-bill-length}
lm(body_mass_g ~ . -bill_length_mm, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.871 -->

```{r minus-island}
lm(body_mass_g ~ . -island, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.874 -->

```{r minus-species}
lm(body_mass_g ~ . -species, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.843 -->

::: {.callout-caution}
# Which variable should be deleted?

Based on the adjusted $R^2$ values, which variable should be deleted from the model. 

*Remember:* You are looking for the model that has a higher adjusted $R^2$ than what you started with!
:::

<!-- delete island -->

<!-- baseline = 0.874 -->

## Models Deleting Two Explanatory Variables

```{r minus-island-year}
lm(body_mass_g ~ . - island - year, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.873 -->

```{r minus-island-sex}
lm(body_mass_g ~ . - island - sex, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.849 -->

```{r minus-island-flipper-length}
lm(body_mass_g ~ . - island - flipper_length_mm, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.861 -->

```{r minus-island-bill-depth}
lm(body_mass_g ~ . - island - bill_depth_mm, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.871 -->

```{r minus-island-bill-length}
lm(body_mass_g ~ . - island - bill_length_mm, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.871 -->

```{r minus-island-species}
lm(body_mass_g ~ . - island - species, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.829 -->

<!-- model missing island is best = 0.871  -->

## Models Deleting Three Explanatory Variables

```{r minus-island-bill-depth-year}
lm(body_mass_g ~ . - island - bill_depth_mm - year, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.873 -->

```{r minus-island-bill-depth-sex}
lm(body_mass_g ~ . - island - bill_depth_mm - sex, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.849 -->

```{r minus-island-bill-depth-flipper-length}
lm(body_mass_g ~ . - island - bill_depth_mm - flipper_length_mm, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.861 -->

```{r minus-island-bill-depth-species}
lm(body_mass_g ~ . - island - bill_depth_mm - species, data = penguins) %>% 
  get_regression_summaries()
```

<!-- 0.871 -->

```{r minus-island-bill-length-bill-depth}
lm(body_mass_g ~ . - island - bill_length_mm - bill_depth_mm, data = penguins) %>% 
  get_regression_summaries()
```

## Models Deleting Four Explanatory Variables

```{r minus-island-year-bill-depth-sex}
lm(body_mass_g ~ . - island - bill_depth_mm - year - sex, data = penguins) %>% 
  get_regression_summaries()
```

```{r minus-island-year-bill-depth-flipper}
lm(body_mass_g ~ . - island - bill_depth_mm - year - flipper_length_mm, data = penguins) %>% 
  get_regression_summaries()
```

```{r minus-island-year-bill-depth-length}
lm(body_mass_g ~ . - island - bill_depth_mm - year - bill_length_mm, data = penguins) %>% 
  get_regression_summaries()
```

```{r minus-island-year-bill-depth-species}
lm(body_mass_g ~ . - island - bill_depth_mm - year - species, data = penguins) %>% 
  get_regression_summaries()
```

::: {.callout-caution}
# Which variable should be deleted?

Based on the adjusted $R^2$ values, which variable should be deleted from the model. 

*Remember:* You are looking for the model that has a higher adjusted $R^2$ than what you started with!
:::


## New Rules

Now, choose the simplest model that is within 1% of the best adjusted $R^2$ you obtained. 

::: {.callout-note}
Note, when I say "simplest model" I mean the model with the fewest variables included. 
:::

::: {.callout-caution}
# What is your best model?

What variables are included in the final model you chose?
:::

